
\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage[pagenumbers]{cvpr} %

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{makecell}
\usepackage[export]{adjustbox}
\usepackage[normalem]{ulem}
\usepackage{textpos}  %


\usepackage{tabulary,multirow,overpic,xcolor,subfloat}
\usepackage[pagebackref=false, breaklinks=true, letterpaper=true, colorlinks,
            citecolor=citecolor, linkcolor=linkcolor, bookmarks=false]{hyperref}
\definecolor{citecolor}{HTML}{0071BC}
\definecolor{linkcolor}{HTML}{ED1C24}

\newcommand{\bd}[1]{\textbf{#1}}
\newcommand{\app}{\raise.17ex\hbox{$\scriptstyle\sim$}}
\newcommand{\ncdot}{{\mkern 0mu\cdot\mkern 0mu}}
\def\x{\times}
\newcolumntype{x}[1]{>{\centering\arraybackslash}p{#1pt}}
\newcolumntype{y}[1]{>{\raggedright\arraybackslash}p{#1pt}}
\newcommand{\dt}[1]{\fontsize{5pt}{0.1em}\selectfont (#1)}
\newlength\savewidth\newcommand\shline{\noalign{\global\savewidth\arrayrulewidth
  \global\arrayrulewidth 1pt}\hline\noalign{\global\arrayrulewidth\savewidth}}
\newcommand{\tablestyle}[2]{\setlength{\tabcolsep}{#1}\renewcommand{\arraystretch}{#2}\centering\footnotesize}
\makeatletter\renewcommand\paragraph{\@startsection{paragraph}{4}{\z@}
  {.5em \@plus1ex \@minus.2ex}{-.5em}{\normalfont\normalsize\bfseries}}\makeatother

\newcommand\blfootnote[1]{\begingroup\renewcommand\thefootnote{}\footnote{#1}\addtocounter{footnote}{-1}\endgroup}

\DeclareMathAlphabet\mathbfcal{OMS}{cmsy}{b}{n}

\newcommand{\std}[1]{\fontsize{5pt}{0.1em}\selectfont #1}
\definecolor{Gray}{gray}{0.5}
\newcommand{\demph}[1]{\textcolor{Gray}{#1}}

\newcommand{\modelname}{Mask2Former\xspace}

\newcommand{\modelnamelong}{Masked-attention Mask Transformer\xspace}
\newcommand{\sota}[0]{state-of-the-art\xspace}

\newcommand{\rg}[1]{{\color{red} R: #1}}
\newcommand{\im}[1]{{\color{magenta} I: #1}}
\newcommand{\rgedit}[1]{{\color{pink}#1}}
\newcommand{\todo}[1]{{\color{red} TODO: #1}}


\usepackage[capitalize]{cleveref}
\crefname{section}{Sec.}{Secs.}
\Crefname{section}{Section}{Sections}
\Crefname{table}{Table}{Tables}
\crefname{table}{Tab.}{Tabs.}

\newcommand{\figref}[1]{Figure~\ref{#1}}
\newcommand{\secref}[1]{Section~\ref{#1}}
\newcommand{\tabref}[1]{Table~\ref{#1}}
\newcommand{\appref}[1]{Appendix~\ref{#1}}


\def\cvprPaperID{*****}
\def\confName{CVPR}
\def\confYear{2022}

\crefname{section}{\S}{\S\S}
\crefname{subsection}{\S}{\S\S}
\crefformat{table}{Table~#2#1#3}
\crefformat{figure}{Figure~#2#1#3}
\crefformat{equation}{Equation~(#2#1#3)}
\crefformat{algorithm}{Algorithm~#2#1#3}
\crefformat{appendix}{Appendix~#2#1#3}

\newcommand{\authorskip}{\hspace{2.5mm}}

\begin{document}





\title{\Large \modelnamelong for Universal Image Segmentation}




\author{
 Bowen Cheng$^{1,2}$\thanks{Work done during an internship at Facebook AI Research.} \authorskip Ishan Misra$^1$ \authorskip Alexander G. Schwing$^2$ \authorskip
 Alexander Kirillov$^1$ \authorskip Rohit Girdhar$^1$ \\
 $^1$Facebook AI Research (FAIR) \authorskip $^2$University of Illinois at Urbana-Champaign (UIUC) \\
{\small \url{https://bowenc0221.github.io/mask2former}}
}

\maketitle

\input{sections/abs}


\input{sections/intro}
\input{sections/relwork}
\input{sections/approach}
\input{sections/expts}
\input{sections/concl}

{
\footnotesize
{\bf \noindent Ethical considerations:} 
While our technical innovations do not appear to have any inherent biases, the models trained with our approach on real-world datasets should undergo ethical review to ensure the predictions do not propagate problematic stereotypes, and the approach is not used for applications including but not limited to illegal surveillance.

\noindent\textbf{Acknowledgments:} Thanks to Nicolas Carion and Xingyi Zhou for helpful feedback. BC and AS are supported in part by NSF \#1718221, 2008387, 2045586, 2106825, MRI \#1725729, NIFA 2020-67021-32799 and Cisco Systems Inc.\ (CG 1377144 - thanks for access to Arcetri).
}

{\small
\bibliographystyle{ieee_fullname}
\bibliography{refs}
}


\clearpage
\appendix
\begin{center}{\bf \Large Appendix}\end{center}\vspace{-2mm}
\renewcommand{\thetable}{\Roman{table}}
\renewcommand{\thefigure}{\Roman{figure}}
\setcounter{table}{0}
\setcounter{figure}{0}
\input{sections/appendix}

\end{document}
