% last updated in April 2002 by Antje Endemann
% Based on CVPR 07 and LNCS, with modifications by DAF, AZ and elle, 2008 and AA, 2010, and CC, 2011; TT, 2014; AAS, 2016

\documentclass[runningheads]{llncs}
\usepackage{graphicx}
\usepackage{amsmath,amssymb} % define this before the line numbering.
%\usepackage{ruler}
\usepackage{color}
\usepackage[width=122mm,left=12mm,paperwidth=146mm,height=193mm,top=12mm,paperheight=217mm]{geometry}


\usepackage[boxed]{algorithm2e}
\usepackage{caption}
\usepackage{mathtools}
\usepackage{amsfonts}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{soul}
\usepackage{color}
\usepackage{xspace}
\usepackage[subrefformat=parens]{subfig}
\usepackage{soul}

\newlength{\soustable}
\setlength{\soustable}{0.3em}

\makeatletter
\DeclareRobustCommand\onedot{\futurelet\@let@token\@onedot}
\def\@onedot{\ifx\@let@token.\else.\null\fi\xspace}
\def\eg{\emph{e.g}\onedot} \def\Eg{\emph{E.g}\onedot}
\def\ie{\emph{i.e}\onedot} \def\Ie{\emph{I.e}\onedot}
\def\cf{\emph{c.f}\onedot} \def\Cf{\emph{C.f}\onedot}
\def\etc{\emph{etc}\onedot}
\def\vs{\emph{vs}\onedot}
\def\wrt{w.r.t\onedot}
\def\etal{\emph{et al}\onedot}
%\def\OURS{T\&G\xspace}
\def\OURS{DeepCluster\xspace}
\newcommand{\ourmethod}{\OURS}
\newcommand{\myarrow}[0]{\small{$\to$}}
\makeatother

\begin{document}
%\renewcommand\thelinenumber{\color[rgb]{0.2,0.5,0.8}\normalfont\sffamily\scriptsize\arabic{linenumber}\color[rgb]{0,0,0}}
%\renewcommand\makeLineNumber {\hss\thelinenumber\ \hspace{6mm} \rlap{\hskip\textwidth\ \hspace{6.5mm}\thelinenumber}}
%\linenumbers
\pagestyle{headings}
\mainmatter
\def\ECCV18SubNumber{1370}  % Insert your submission number here


\title{Deep Clustering for Unsupervised Learning \\  of Visual Features} % Replace with your title

\titlerunning{Deep Clustering for Unsupervised Learning of Visual Features}

\authorrunning{Mathilde Caron \etal}

\author{Mathilde Caron \and Piotr Bojanowski \and Armand Joulin \and Matthijs Douze}
\institute{Facebook AI Research}


\maketitle

\begin{abstract}
  Clustering is a class of unsupervised learning methods that has been extensively applied and studied in computer vision.
  Little work has been done to adapt it to the end-to-end training of visual features on large scale datasets.
  In this work, we present \OURS, a clustering method that jointly learns the parameters of a neural network and the cluster assignments
  of the resulting features.
  \OURS iteratively groups the features with a standard clustering algorithm, $k$-means, and uses the subsequent assignments as supervision
  to update the weights of the network.
  We apply \OURS to the unsupervised training of convolutional neural networks on large datasets like ImageNet and YFCC100M.
  The resulting model outperforms the current state of the art by a significant margin on all the standard benchmarks.
%  We validate this finding by experimenting with different training sets and architectures.
\keywords{unsupervised learning, clustering}
\end{abstract}

\section{Introduction}

\input{intro.tex}




\section{Related Work}

\input{related.tex}



\section{Method}

\input{method.tex}

\section{Experiments}

\input{results.tex}

\input{discussion.tex}
\section{Conclusion}

In this paper, we propose a scalable clustering approach for the unsupervised learning of convnets.
It iterates between clustering  with $k$-means the features produced by the convnet and
updating its weights by predicting the cluster assignments as pseudo-labels in a discriminative loss.
If trained on large dataset like ImageNet or YFCC100M, it achieves performance that are significantly
better than the previous state-of-the-art on every standard transfer task.
Our approach makes little assumption about the inputs, and does not require much domain specific knowledge, making it a
good candidate to learn deep representations specific to domains where annotations are scarce.

\paragraph{Acknowledgement.}
We thank Alexandre Sablayrolles and the rest of the FAIR team for their feedback and fruitful discussion around this paper.
We would like to particularly thank Ishan Misra for spotting an error in our evaluation setting of Table~\ref{tab:linear}.

\clearpage

\bibliographystyle{splncs}
\bibliography{egbib}

\clearpage
\input{supp.tex}

\end{document}
