%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TODOs:
% * References: GraphStar, GraphVite, GaAN
% * Observe the differences in dataset stats. what kind of graphs does cluster-GCN fail? skewed?
%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass{article} % For LaTeX2e
\usepackage{iclr2020_conference,times}

% Optional math commands from https://github.com/goodfeli/dlbook_notation.
%%\input{math_commands.tex}

\usepackage{hyperref}
\usepackage{url}

\usepackage{tikz}
\usetikzlibrary{fit,backgrounds,arrows,decorations.markings,arrows.meta}

\usepackage{graphicx}
\usepackage{bm}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts,amsmath}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{multirow}
\usepackage{algorithm}% http://ctan.org/pkg/algorithms
\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
\usepackage{mathtools}
\usepackage[inline]{enumitem}
\usepackage{pgfplots}
\usepackage{relsize}

\usepgfplotslibrary{groupplots}

% \pgfplotsset{
%         % define the layers you need.
%         % (Don't forget to add `main' somewhere in that list!!)
%         layers/my layer set/.define layer set={
%             background,
%             main,
%             foreground
%         }{
%             % you could state styles here which should be moved to
%             % corresponding layers, but that is not necessary here.
%             % That is why wo don't state anything here
%         },
%         % activate the newly created layer set
%         set layers=my layer set,
%     }
\usetikzlibrary{matrix}
\usepackage{amsthm,bbm}

% \pgfplotsset{compat=newest}
% \usepgfplotslibrary{groupplots}
% \usepgfplotslibrary{dateplot}

% \usepackage{adjustbox}
%%\usepackage{subfigure}
\usepackage{subcaption}
\usepackage{ctable}

\usepackage[symbol]{footmisc}

\renewcommand{\thefootnote}{\fnsymbol{footnote}}
%\usepackage[dvipsnames]{xcolor}

\DeclarePairedDelimiterX\set[1]\lbrace\rbrace{\def\given{\;\delimsize\vert\;}#1}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}


\newcommand{\trans}{\mathsf{T}}

\newcommand{\graphsaint}{{\fontfamily{lmtt}\selectfont GraphSAINT}}
\newcommand{\saint}{{\fontfamily{lmtt}\selectfont SAINT}}

\newcommand{\sample}{{\fontfamily{lmtt}\selectfont SAMPLE}}
\newcommand{\G}[1][]{
    \ifthenelse{\equal{#1}{}}
    {\mathcal{G}}
    {\mathcal{G}}_{#1}}
\newcommand{\V}[1][]{
    \ifthenelse{\equal{#1}{}}
    {\mathcal{V}}
    {\mathcal{V}_{#1}}}
\newcommand{\E}[1][]{
    \ifthenelse{\equal{#1}{}}
    {\mathcal{E}}
    {\mathcal{E}_{#1}}}
\newcommand{\X}[1][]{
    \ifthenelse{\equal{#1}{}}
    {\bm{X}}
    {\bm{X}^{\paren{#1}}}}
\newcommand{\Anorm}[1][]{
    \ifthenelse{\equal{#1}{}}
    {\widetilde{\bm{A}}}
    {\widetilde{\bm{A}}_{#1}}}
\newcommand{\xvl}[2]{\bm{x}_{#1}^{(#2)}}
\newcommand{\W}[1][]{
    \ifthenelse{\equal{#1}{}}
    {\bm{W}}
    {\bm{W}^{\paren{#1}}}}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}



\newcommand{\size}[1]{\left\lvert #1 \right\rvert}
\newcommand{\paren}[1]{\left( #1 \right)}


\tikzset{
    vertex/.style = {
        circle,
        fill            = black,
        outer sep = 2pt,
        inner sep = 1pt,
    }
}




\title{\graphsaint: \underline{Graph Sa}mpling Based  \underline{In}ductive Learning Me\underline{t}hod}

% Authors must not appear in the submitted version. They should be hidden
% as long as the \iclrfinalcopy macro remains commented out below.
% Non-anonymous submissions will be rejected without review.



% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to \LaTeX{} to determine where to break
% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
% puts 3 of 4 authors names on the first line, and the last on the second
% line, try using \AND instead of \And before the third author name.
\author{Hanqing Zeng\thanks{Equal contribution}\\
University of Southern California\\
\texttt{zengh@usc.edu}\\
\And
Hongkuan Zhou\footnotemark[1]\\
University of Southern California\\
\texttt{hongkuaz@usc.edu}\\
\And
Ajitesh Srivastava\\
University of Southern California\\
\texttt{ajiteshs@usc.edu}
\And
Rajgopal Kannan\\
US Army Research Lab\\
\texttt{rajgopal.kannan.civ@mail.mil}
\And
Viktor Prasanna\\
University of Southern California\\
\texttt{prasanna@usc.edu}}
\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\iclrfinalcopy % Uncomment for camera-ready version, but NOT for submission.
\begin{document}


\maketitle

\begin{abstract}
Graph Convolutional Networks (GCNs) are powerful models for learning representations of attributed graphs. %, thus facilitating classification and clustering tasks. 
To scale GCNs to large graphs, state-of-the-art methods use various \emph{layer sampling} techniques to alleviate the ``neighbor explosion'' problem during minibatch training. 
We propose {\graphsaint}, a \emph{graph sampling} based inductive learning method that improves training efficiency and accuracy in a fundamentally different way. 
By changing perspective, {\graphsaint} constructs minibatches by sampling the training graph, rather than the nodes or edges across GCN layers. 
%We change the viewpoint from layer sampling to graph sampling. The minibatch construction is decoupled from the forward/backward propagation in training. 
Each iteration, a complete GCN is built from the properly sampled subgraph. Thus, we ensure fixed number of well-connected nodes in all layers. 
We further propose normalization technique to eliminate bias, and sampling algorithms for variance reduction. 
Importantly, we can decouple the sampling from the forward and backward propagation, and extend {\graphsaint} with many architecture variants (e.g., graph attention, jumping connection).  
%lightweight sampling process that can be carried out even before the training starts; nodes across layers are well connected. 
{\graphsaint} demonstrates superior performance in both accuracy and training time on five large graphs, and achieves new state-of-the-art F1 scores for PPI (0.995) and Reddit (0.970). 
\end{abstract}
\input{1_intro.tex}
\input{2_related.tex}
\input{3_method.tex}
\input{4_discussion.tex}
\input{5_exp.tex}
\input{6_conclusion.tex}

\subsubsection*{acknowledgement}
This material is based on work supported
by the Defense Advanced Research Projects Agency (DARPA) under Contract Number FA8750-17-C-0086 and National Science Foundation (NSF) under Contract Numbers CCF-1919289 and OAC-1911229. Any opinions, findings and conclusions or recommendations
expressed in this material are those of the authors and do not necessarily reflect the views of DARPA or NSF.

\bibliography{citation}
\bibliographystyle{iclr2020_conference}

\appendix
\input{7_appendix.tex}

\end{document}
