\begin{abstract}





Large language models (LLMs) have transformed numerous AI applications. \textit{On-device} LLM is becoming increasingly important: running LLMs locally on edge devices can reduce the cloud computing cost and protect users' privacy. However, the astronomical model size and the limited hardware resource pose significant deployment challenges. We propose \method (\methodshort), a hardware-friendly approach for LLM low-bit weight-only quantization. \methodshort finds that not all weights in an LLM are equally important. Protecting \textit{only 1\%} salient weights can greatly reduce quantization error. To identify salient weight channels, we should refer to the activation distribution, not weights. 
To avoid the hardware-inefficient mix-precision quantization, we mathematically derive that scaling up the salient channels can reduce the quantization error.
\methodshort employs an equivalent transformation to scale the salient weight channels to protect them. The scale is determined by collecting the activation statistics offline.
\methodshort does not rely on any backpropagation or reconstruction, so it generalizes to different domains and modalities without overfitting the calibration set. \methodshort outperforms existing work on various language modeling and domain-specific benchmarks (coding and math). Thanks to better generalization, it achieves excellent quantization performance for \emph{instruction-tuned} LMs and, for the first time, \emph{multi-modal} LMs. 
Alongside AWQ, we implement \system, an efficient and flexible inference framework tailored for 4-bit on-device LLM/VLMs. With kernel fusion and platform-aware weight packing, \system offers more than \textbf{3}$\times$ speedup over the Huggingface FP16 implementation on both desktop and mobile GPUs. It also democratizes the deployment of the 70B Llama-2 model on mobile GPUs. 

\end{abstract}


