\documentclass{article} %

\usepackage[utf8]{inputenc} %
\usepackage[T1]{fontenc}    %
\usepackage[hyperfootnotes=false]{hyperref}       %
\usepackage{url}            %
\usepackage{booktabs}       %
\usepackage{amsmath,amsfonts,amsthm}       %
\usepackage{mathtools}      %
\usepackage{nicefrac}       %
\usepackage{microtype}      %
\usepackage{subcaption}
\usepackage{bbm}
\usepackage{multirow}
\usepackage[inline]{enumitem}
\usepackage{diagbox}
\usepackage{pifont}
\usepackage[capitalise]{cleveref}  %
\usepackage{comment}
\usepackage{etoolbox}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage[font=small]{caption}


\graphicspath{{figures/}{../figures/}}
\usepackage{import}
\usepackage{subfiles}
\providecommand{\main}{.}

\newtoggle{arxiv}
\toggletrue{arxiv}


\newcommand\blfootnote[1]{%
  \begingroup
  \renewcommand\thefootnote{}\footnote{#1}%
  \addtocounter{footnote}{-1}%
  \endgroup
}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}[section]
\newtheorem{corollary}[lemma]{Corollary}
\newtheorem{observation}[lemma]{Observation}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}{Definition}
\newtheorem{remark}[lemma]{Remark}
\newtheorem{claim}{Claim}
\newtheorem{innercustomthm}{Theorem}
\newenvironment{customthm}[1]
{\renewcommand\theinnercustomthm{#1}\innercustomthm}
{\endinnercustomthm}

\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Rn}[1]{(-\infty, #1]}
\renewcommand{\d}{\mathop{}\!\mathrm{d}}
\newcommand{\dd}{\mathop{}\!\mathrm{d}} %
\newcommand{\partialt}{\partial t} %
\newcommand{\ppt}{\frac{\partial}{\partial t}}
\newcommand{\ddt}{\frac{d}{d t}}
\newcommand{\defeq}{\coloneqq}
\newcommand{\norm}[1]{\left\|{#1}\right\|} %
\providecommand{\abs}[1]{\left\lvert#1\right\rvert}
\DeclareMathOperator*{\diag}{diag}
\DeclareMathOperator{\proj}{proj}
\DeclareMathOperator{\rec}{rec}
\DeclareMathOperator{\coef}{coef}
\DeclareMathOperator{\hippo}{hippo}

\newcommand*\samethanks[1][\value{footnote}]{\footnotemark[#1]}


\usepackage[numbers,sort]{natbib}
\setlength{\textwidth}{6.5in}
\setlength{\textheight}{9in}
\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\topmargin}{-0.5in}
\newlength{\defbaselineskip}
\setlength{\defbaselineskip}{\baselineskip}
\setlength{\marginparwidth}{0.8in}


\title{HiPPO: Recurrent Memory with Optimal Polynomial Projections}
\usepackage{authblk}
\author[$\dagger$]{Albert Gu\thanks{Equal contribution. Order determined by coin flip.}}
\author[$\dagger$]{Tri Dao\samethanks}
\author[$\dagger$]{Stefano Ermon}
\author[$\ddagger$]{Atri Rudra}
\author[$\dagger$]{Christopher R{\'e}}
\affil[$\dagger$]{Department of Computer Science, Stanford University}
\affil[$\ddagger$]{Department of Computer Science and Engineering, University at Buffalo, SUNY\vspace{4pt}}
\affil[ ]{{\texttt{\{albertgu,trid\}@stanford.edu}, \texttt{ermon@cs.stanford.edu}, \texttt{atri@buffalo.edu}, \texttt{chrismre@cs.stanford.edu}}}

\begin{document}

\maketitle

\begin{abstract}
  A central problem in learning from sequential data is representing cumulative
  history in an incremental fashion as more data is processed.
  We introduce a general framework (HiPPO) for the online compression of continuous signals and discrete time series
  by projection onto polynomial bases.
  Given a measure that specifies the importance of each time step in the past,
  HiPPO produces an optimal solution to a natural \emph{online function approximation} problem.
  As special cases, our framework yields a short derivation of the recent Legendre Memory Unit (LMU) from first principles,
  and generalizes the ubiquitous gating mechanism of recurrent neural networks such as GRUs.
  This formal framework yields a new memory update mechanism (HiPPO-LegS) that scales
  through time to remember all history, avoiding priors on the timescale.
  HiPPO-LegS enjoys the theoretical benefits of timescale robustness, fast updates,
  and bounded gradients.
  By incorporating the memory dynamics into recurrent neural networks, HiPPO RNNs can
  empirically capture complex temporal dependencies.
  On the benchmark permuted MNIST dataset, HiPPO-LegS sets a new
  state-of-the-art accuracy of 98.3\%.
  Finally, on a novel trajectory classification task testing robustness to out-of-distribution timescales and missing data, HiPPO-LegS outperforms RNN and neural ODE baselines by 25-40\% accuracy.
\end{abstract}


\input{src/intro.tex}

\input{src/framework.tex}

\input{src/legs.tex}

\input{src/experiments.tex}

\input{src/conclusion.tex}


\subsubsection*{Acknowledgments}

We thank
Avner May, Mayee Chen, Dan Fu,
Aditya Grover,
and Daniel L\'{e}vy
for their helpful feedback.
We gratefully acknowledge the support of DARPA under Nos.\ FA87501720095 (D3M),
FA86501827865 (SDH), and FA86501827882 (ASED); NIH under No.\ U54EB020405
(Mobilize), NSF under Nos.\ CCF1763315 (Beyond Sparsity), CCF1563078 (Volume to
Velocity), and 1937301 (RTML); ONR under No.\ N000141712266 (Unifying Weak
Supervision); the Moore Foundation, NXP, Xilinx, LETI-CEA, Intel, IBM,
Microsoft, NEC, Toshiba, TSMC, ARM, Hitachi, BASF, Accenture, Ericsson,
Qualcomm, Analog Devices, the Okawa Foundation, American Family Insurance,
Google Cloud, Stanford HAI AWS cloud credit, Swiss Re, and members of the Stanford DAWN project: Teradata,
Facebook, Google, Ant Financial, NEC, VMWare, and Infosys.
The U.S.\ Government is authorized to reproduce and distribute reprints for
Governmental purposes notwithstanding any copyright notation thereon.
Any opinions, findings, and conclusions or recommendations expressed in this
material are those of the authors and do not necessarily reflect the views,
policies, or endorsements, either expressed or implied, of DARPA, NIH, ONR, or
the U.S.\ Government.
Atri Rudra’s research is supported by NSF grant CCF-1763481.


\bibliography{hippo}
\bibliographystyle{plainnat}

\appendix

\newpage

\subfile{src/rw}


\subfile{src/background}

\subfile{src/framework_details}

\subfile{src/derivations}

\subfile{src/theory}

\subfile{src/experiment_details}







\end{document}
