
\section{Introduction}
% Introduce the problem that you are trying to solve and write 1-2 paragraphs for each of the sub-sections. Provide a high-level summary in this section instead of detailed descriptions.

% \paragraph{Problem Definition.}


In many scientific domains such as climate science, neurology, and epidemiology, low-level sensor measurements generate high-dimensional observational data. These data are naturally represented as gridded time series, with interactions that evolve over both space and time. Discovering causal relationships from spatiotemporal  data is of great  scientific importance.  It allows researchers to predict future states, intervene in harmful trends, and develop new insights into the underlying mechanisms. For example, in climate science, the study of teleconnections \citep{liu2023teleconnections}, the interactions between regions thousands of kilometers away, is important to understanding how climate events in one part of the world may affect weather patterns in distant locations. 
% Similarly, in neurology, the influences between different brain regions reveal the brain's organization during cognitive tasks and in various neurological conditions \citep{friston2009causal,arab2023whole}.
\input{figures/spacy_overview}
\input{figures/theory_compare}

Despite the plethora of work on causal structure learning from time series data \citep{granger1969investigating, hyvarinen2010estimation, runge2020discovering, tank2021neural, gong2022rhino, cheng2023cutsneuralcausaldiscovery}, they face significant challenges for  high-dimensional spatiotemporal data. A primary reason is scalability. The high dimensionality of gridded data makes it difficult for many of these techniques, especially those relying on conditional independence tests, to scale effectively \citep{glymour2019review}. Additionally, spatially proximate points often exhibit redundant and highly correlated time series. Conditioning on nearby correlated points can obscure true causal relationships between distant locations, reducing the statistical power of conditional independence tests and leading to inaccurate results \citep{xavi2022teleconnection}.

Recent advances in spatiotemporal causal discovery typically follow a two-stage process: first, dimensionality reduction is applied to extract a small number of latent time series from the original grid of time series; then, causal discovery is performed on these low-dimensional representations. Examples of this approach include \citet{xavi2022teleconnection} and \citet{Falasca_2024}. However, these two-stage approaches perform dimensionality reduction independent of the causal structure, potentially leading to latent representations that obscure the relationships among causally relevant entities. Another important line of research is causal representation learning from time series data \citep{scholkopf2021toward}. While approaches like \citet{yao2021learning, yao2022temporally, chencaring} infer latent time series from high-dimensional data, they do not incorporate spatial priors, making them less suitable for spatiotemporal causal discovery. \citet{brouillard2024causal} learn to map each  time series to a latent variable under the single-parent assumption, that is, each observed variable is influenced by only one latent variable. However, this assumption can be overly restrictive in spatiotemporal systems where observed variables are often influenced by multiple interacting latent factors (e.g., atmospheric patterns, ocean currents) that jointly drive their behavior.
% However, it assumes that each observed variable has only one latent parent.

% \input{figures/intro}

We present SPAtiotemporal Causal DiscoverY (SPACY), a novel causal representation learning framework for spatiaotemporal data, to address these limitations (Figure \ref{fig:spacy_overview}). To model spatial variability on the grid, we introduce spatial factors, parametrized by spatial kernel functions such as Radial Basis Functions (RBFs). These spatial factors aggregate proximate grid points and map the observed variables to their corresponding latent time series. Since SPACY performs causal discovery on these lower-dimensional representations, SPACY is naturally scalable. We derive an evidence lower bound to learn the spatial factors, latent time series as well as the causal graph. Our approach jointly infers both the latent time series and the underlying causal graph in an \textit{end-to-end} manner.

% We also prove the identifiability of our framework for a continuous spatial domain with an infinite resolution. We demonstrate that the overdetermined nature of the system allows us to uniquely recover the spatial factors and latent time series (up to permutation and scaling) with minimal assumptions about the latent generative process. Notably, compared to previous works, our framework can handle instantaneous edges while also allowing observed variables to be associated with multiple latent parents.

We also prove the identifiability of our framework for a continuous spatial domain with an infinite resolution. \change{We show that when the observations are a nonlinear, invertible function of the product of latent series and spatial factors, we can leverage the overdetermined structure of the system to recover spatial factors and latent time series (up to permutation and scaling)}. Notably, compared to previous works, our framework can handle instantaneous edges while also allowing observed variables to be associated with multiple latent parents.



Our main contributions can be summarized as follows.
\begin{enumerate}[noitemsep, topsep=0pt]
    \item We introduce SPACY, a novel variational inference-based spatiotemporal causal discovery framework that jointly infers latent time series and the underlying causal graph.
    \item Theoretically, we analyze the identifiability of our system for continuous spatial domains. We show that the latent factors are identifiable up to permutation and scaling under both linear and nonlinear invertible mappings between the latent and observed variables.
    \item Experimentally, we demonstrate the strong performance of our method on both synthetic and real-world datasets. SPACY accurately recovers causal links in scenarios involving both linear and nonlinear interactions, including settings where existing methods fail. SPACY is also highly scalable, and can infer causal links from high-dimensional grids (upto $250\times 250$).
\end{enumerate}

% Spatio-temporal data is a collection of time series that contains location attribute, which describes patterns of both chronological and spatial relationships. Identifying the causal relations between variables and their interactions through time and space is essential for various domains such as epidemiology, economics, and climatology\citep{xavi2022teleconnection}\citep{ebert2014climate}. 

% The task of Spatio-temporal causal discovery can be challenging for several reasons: (1) The high-dimensional nature of Spatio-temporal data; (2) The diversity in spatial assumptions across different variates; (3) proximity issue: closely located data points might show spurious relationships. For example, to determine the causes of unusual rainfall in California over the past year, one might look at the climate data of West Coast North America, which consists of time series across 1,300 miles (i.e. high dimensional). The climate data would consist of time series from different variates, for instance, precipitation and sea surface temperature \citep{Falasca_2024} (i.e. different variates). Lastly, one may need to pay attention to spatial proximity: while rain in Irvine may correlate strongly with precipitation in Los Angeles, such correlations might not reflect true causal relationships due to their proximity (i.e. proximity issue). Moreover, these correlations can overshadow the true causal relationships, complicating the detection of causal links during analysis.

% Due to these inherent challenges, the task of spatio-temporal causal discovery has yet to be addressed satisfactorily. Numerous studies have been devoted to spatio-temporal causal discovery, with approaches that combine traditional dimensionality reduction methods with conditional independence tests. \citet{xavi2022teleconnection} introduced a framework combining Varimax rotated PCA with PCMCI \citep{runge2018pcmci, runge2020pcmci+} to deduce lagged effects in gridded time series. However, due to the non-differentiability, PCMCI's reliance on the validity of dimensionality reduction and linear causal assumptions limits its effectiveness in capturing complex, real-world interactions. Correlation and distance-based dimensionality reduction have also been explored in recent work \cite{Falasca_2024}, which utilized a stronger notion in the spatial attribute. Despite this advancement, such methods still depend heavily on the premise of dimensionality reduction and face increased computational demands due to the need to calculate correlations between time series.

% The alternative line of research would be performing causal discovery utilizing a differentiable model, which interconnects the spatial assumptions with the causal discovery process. Deep learning methods like the Spatial-Temporal Causal Discovery Framework (STCD) \citet{sheth2022stcd} have been proposed to overcome Mapped-PCMCI's challenges. Nonetheless, STCD faces its limitations, such as its dependence on strong scenario-specific spatial assumptions and the need for an additional adjacency matrix to represent grid node connectivity, potentially complicating its application to diverse datasets.

% To address these limitations, we introduce SPACY (SPAtiotemporal Causal discoverY), a novel framework for spatiotemporal causal discovery based on variational inference. SPACY leverages a variational autoencoder to model latent time series and their causal relationships, integrating spatial information directly into the learning process. Unlike traditional approaches, our method does not require dimensionality reduction or strong linear assumptions, allowing it to capture both nonlinear and lagged effects. Additionally, SPACY is scalable to large datasets and can model complex, history-dependent noise, making it a powerful tool for causal discovery in real-world spatiotemporal data.