\section{Introduction}

Deep convolutional neural networks (CNNs) have shown promising performances on various computer vision problems such as image classification~\cite{ImageNet,alexnet,resnet}, object detection~\cite{fasterrcnn,liu2016ssd}, semantic segmentation~\cite{chen2018deeplab,Long_2015_CVPR}, and video analysis~\cite{nam2016learning,simonyan2014two}. 
To further improve the training efficiency and performance, a number of training strategies have been proposed, including data augmentation~\cite{alexnet} and regularization {techniques}~\cite{Dropout,stochasticdepth,szegedy2016rethinking_labelsm}.

In particular, to prevent a CNN from focusing too much on a small set of intermediate activations or on a small region on input images, random feature removal regularizations have been proposed. Examples include dropout~\cite{Dropout} for randomly dropping hidden activations and regional dropout~\cite{devries2017cutout,zhong2017randomerase,singh2017hide,ghiasi2018dropblock,choe2019attention} for erasing random regions on the input.
Researchers have shown that the feature removal strategies improve generalization and localization by letting a model attend not only to the most discriminative parts of objects, but rather to the entire object region~\cite{singh2017hide,ghiasi2018dropblock}.


\input{tables/teaser.tex}



While regional dropout strategies have shown improvements of classification and localization performances to a certain degree, deleted regions are usually zeroed-out~\cite{devries2017cutout,singh2017hide} or filled with random noise~\cite{zhong2017randomerase}, greatly reducing the proportion of informative pixels on training images.
We recognize this as a severe conceptual limitation as CNNs are generally data hungry~\cite{InstagramNet}.
How can we maximally utilize the deleted regions, while taking advantage of better generalization and localization using regional dropout?

We address the above question by introducing an augmentation strategy \textbf{CutMix}. 
Instead of simply removing pixels, we replace the removed regions with a patch from another image (See Table~\ref{tab:teaser}).
The ground truth labels are also mixed proportionally to the number of pixels of combined images.
{CutMix now enjoys the property that there is no uninformative pixel during training, making training efficient, while retaining the advantages of regional dropout to attend to non-discriminative parts of objects.}
The added patches further enhance localization ability by requiring the model to identify the object from a partial view.
The training and inference budgets remain the same.

CutMix shares similarity with Mixup~\cite{zhang2017mixup} which mixes two samples by interpolating both the image and labels.  
{While certainly improving classification performance, Mixup samples tend to be unnatural (See the mixed image in Table~\ref{tab:teaser}).
CutMix overcomes the problem by replacing the image region with a patch from another training image.}


Table~\ref{tab:teaser} gives an overview of Mixup~\cite{zhang2017mixup}, Cutout~\cite{devries2017cutout}, and CutMix on image classification, weakly supervised localization, and transfer learning to object detection methods.
Although Mixup and Cutout enhance ImageNet classification, they decrease the ImageNet localization or object detection performances. 
On the other hand, CutMix consistently achieves significant enhancements across three tasks.
%, proving its superior classification and localization ability beyond the baseline and other augmentation methods.

We present extensive evaluations of CutMix on various CNN architectures, datasets, and tasks. 
Summarizing the key results, CutMix has significantly improved the accuracy of a baseline classifier on CIFAR-100 and has obtained the state-of-the-art top-1 error $14.47\%$. On ImageNet~\cite{ImageNet}, applying CutMix to ResNet-50 and ResNet-101~\cite{resnet} has improved the classification accuracy by $+2.28\%$ and $+1.70\%$, respectively. 
On the localization front, CutMix improves the performance of the weakly-supervised object localization (WSOL) task on CUB200-2011~\cite{wah2011cub} and ImageNet~\cite{ImageNet} by {$+5.4\%$} and $+0.9\%$, respectively. 
The superior localization capability is further evidenced by fine-tuning a detector and an image caption generator on CutMix-ImageNet-pretrained models; the CutMix pretraining has improved the overall detection performances on Pascal VOC~\cite{pascalvoc} by $+1$ mAP and image captioning performance on MS-COCO~\cite{lin2014microsoft} by $+2$ BLEU scores.
{CutMix also enhances the model robustness and alleviates the over-confidence issue \cite{hendrycks2016baseline, liang2017odin} of deep networks.}
