\section{Related Work}

Data augmentation has played a central role in the training of deep vision models. On natural images, horizontal flips and random cropping or translations of the images are commonly used in classification and detection models~\cite{WRN2016,krizhevsky2012imagenet,girshick2018detectron}.  On MNIST, elastic distortions across scale, position, and orientation have been applied to achieve impressive results~\cite{simard2003best,ciregan2012multi,wan2013regularization,sato2015apac}. While previous examples augment the data while keeping it in the training set distribution, operations that do the opposite can also be effective in increasing generalization. Some methods randomly erase or add noise to patches of images for increased validation accuracy~\cite{cutout2017,zhong2017random}, robustness~\cite{szegedy2013intriguing,yin2019afourier,ford2019adversarial}, or both~\cite{lopes2019improving}.  Mixup~\cite{zhang2017mixup} is a particularly effective augmentation method on CIFAR-10 and ImageNet, where the neural network is trained on convex combinations of images and their corresponding labels. Object-centric cropping is commonly used for object detection tasks \cite{liu2016ssd}, whereas \cite{dwibedi2017cut} adds new objects on training images by cut-and-paste.

Moving away from individual operations to augment data, other work has focused on finding optimal strategies for combining different operations. For example, Smart Augmentation learns a network that merges two or more samples from the same class to generate new data~\cite{lemley2017smart}. Tran et al. generate augmented data via a Bayesian approach, based on the distribution learned from the training set~\cite{tran2017bayesian}. DeVries et al. use transformations (e.g. noise, interpolations and extrapolations) in the learned feature space to augment data~\cite{devries2017dataset}. Furthermore, generative adversarial networks (GAN) have been used to choose optimal sequences of data augmentation operations\cite{ratner2017learning}. GANs have also been used to generate training data directly~\cite{perez2017effectiveness,mun2017generative,zhu2017data,antoniou2017data,sixt2016rendergan}, however this approach does not seem to be as beneficial as learning sequences of data augmentation operations that are pre-defined~\cite{ravuri2019classification}.   

Another approach to learning data augmentation strategies from data is AutoAugment~\cite{cubuk2018autoaugment}, which originally used reinforcement learning to choose a sequence of operations as well as their probability of application and magnitude. Application of AutoAugment policies involves stochasticity at multiple levels: 1) for every image in every minibatch, a sub-policy is chosen with uniform probability. 2) operations in each sub-policy has an associated probability of application. 3) Some operations have stochasticity over direction. For example, an image can be rotated clockwise or counter-clockwise. The layers of stochasticity increase the amount of diversity that the network is trained on, which in turn was found to significantly improve generalization on many datasets. More recently, several papers used the AutoAugment search space and formalism with improved optimization algorithms to find AutoAugment policies more efficiently~\cite{ho2019population, lim2019fast}. Although the time it takes to search for policies has been reduced significantly, having to implement these methods in a separate search phase reduces the applicability of AutoAugment. For this reason, this work aims to eliminate the search phase on a separate proxy task completely.    

Some of the developments in RandAugment were inspired by the recent improvements to searching over data augmentation policies. For example, Population Based Augmentation (PBA) \cite{ho2019population} found that the optimal magnitude of augmentations increased during the course of training, which inspired us to not search over optimal magnitudes for each transformation but have a fixed magnitude schedule, which we discuss in detail in Section~\ref{sec:methods}. Furthermore, authors of Fast AutoAugment~\cite{lim2019fast} found that a data augmentation policy that is trained for density matching leads to improved generalization accuracy, which inspired our first order differentiable term for improving augmentation (see Section~\ref{first_order_term}). 
