\label{sec:problem}
\begin{comment}
\todo{
  \begin{enumerate}
    \item we are interested in deep learning methods for object classification and segmentation on 3d point sets.
    \item we solve the two problems in a unified framework. in this framework, we use neural network to learn a function $f$ that takes a subset of points in $\R^3$ as input, and predicts the object recognition or per-point segmentation.  
  \end{enumerate}
}
\end{comment}

We design a deep learning framework that directly consumes unordered point sets as inputs. %, while most recent state-of-the-art methods convert them to volumetric grids \cite{wu20153d,wang2015voting,qi2016volumetric} or projecting them to images \cite{su15mvcnn,qi2016volumetric}. % Furthermore, our proposed network is a unified framework for a various of applications including object classification, object-level part segmentation and scene semantic parsing.
A point cloud is represented as a set of 3D points $\{P_i| \ i=1,...,n\}$, where each point $P_i$ is a vector of its $(x,y,z)$ coordinate plus extra feature channels such as color, normal etc. For simplicity and clarity, unless otherwise noted, we only use the $(x,y,z)$ coordinate as our point's channels. % Our network directly consumes the point cloud 

For the object classification task, the input point cloud is either directly sampled from a shape or pre-segmented from a scene point cloud. Our proposed deep network %first generate global descriptors for the whole input point cloud and then 
outputs $k$ scores for all the $k$ candidate classes.
For semantic segmentation, the input can be a single object for part region segmentation, or a sub-volume from a 3D scene for object region segmentation. Our model will output $n \times m$ scores for each of the $n$ points and each of the $m$ semantic sub-categories.

