\section{Weakly-supervised Object Localization}
\label{appendix:wsol}




We describe the training and evaluation procedure of weakly-supervised object localization in detail.

\noindent\textbf{Network modification: }
Basically weakly-supervised object localization (WSOL) has the same training strategy as image classification does. 
Training WSOL is starting from ImageNet-pretrained model. 
From the base network structures, VGG-16 and ResNet-50~\cite{resnet}, WSOL takes larger spatial size of feature map $14 \times 14$ whereas the original models has $7 \times 7$.
For VGG network, we utilize VGG-GAP, which is a modified VGG-16 introduced in~\cite{zhou2016CAM}.
For ResNet-50, we modified the final residual block (\texttt{layer4}) to have no stride ($=1$), which originally has stride $2$. 

Since the network is modified and the target dataset could be different from ImageNet~\cite{ImageNet}, the last fully-connected layer is randomly initialized with the final output dimension of $200$ and $1000$ for CUB200-2011~\cite{wah2011cub} and ImageNet, respectively.


\noindent\textbf{Input image transformation: }
For fair comparison, we used the same data augmentation strategy except Mixup, Cutout, and CutMix as the state-of-the-art WSOL methods do~\cite{singh2017hide,zhang2018ACOL}.
In training, the input image is resized to $256\times256$ size and randomly cropped $224 \times 224$ size images are used to train network.
In testing, the input image is resized to $256\times256$, cropped at center with $224\times224$ size and used to validate the network, which called single crop strategy. 


\noindent\textbf{Estimating bounding box: }
We utilize class activation mapping (CAM)~\cite{zhou2016CAM} to estimate the bounding box of an object. 
First we compute CAM of an image, and next, we decide the foreground region of the image by binarizing the CAM with a specific threshold. 
The region with intensity over the threshold is set to 1, otherwise to 0. 
We use the threshold as a specific rate $\sigma$ of the maximum intensity of the CAM. 
We set $\sigma$ to $0.15$ for all our experiments. 
From the binarized foreground map, the tightest box which can cover the largest connected region in the foreground map is selected to the bounding box for WSOL.


\noindent\textbf{Evaluation metric: }
To measure the localization accuracy of models, we report top-1 localization accuracy (Loc), which is used for ImageNet localization challenge~\cite{ImageNet}. 
For top-1 localization accuracy, intersection-over-union (IoU) between the estimated bounding box and ground truth position is larger than $0.5$, and, at the same time, the estimated class label should be correct.
Otherwise, top-1 localization accuracy treats the estimation was wrong.


\subsection{CUB200-2011}
CUB-200-2011 dataset~\cite{wah2011cub} contains over 11 K images with 200 categories of birds.
We set the number of training epochs to $600$. 
For ResNet-50, the learning rate for the last fully-connected layer and the other were set to $0.01$ and $0.001$, respectively.
For VGG network, the learning rate for the last fully-connected layer and the other were set to $0.001$ and $0.0001$, respectively.
The learning rate is decaying by the factor of $0.1$ at every $150$ epochs.
We used SGD optimizer, and the minibatch size, momentum, weight decay were set to $32$, $0.9$, and $0.0001$. 

\subsection{ImageNet dataset}
ImageNet-1K~\cite{ImageNet} is a large-scale dataset for general objects consisting of 13 M training samples and 50 K validation samples.
We set the number of training epochs to $20$. 
The learning rate for the last fully-connected layer and the other were set to $0.1$ and $0.01$, respectively.
The learning rate is decaying by the factor of $0.1$ at every $6$ epochs.
We used SGD optimizer, and the minibatch size, momentum, weight decay were set to $256$, $0.9$, and $0.0001$.


