

%\documentclass{article}
\documentclass[10pt,reqno]{amsart}
\usepackage{amsmath,amsthm,amssymb}
\usepackage[utf8]{inputenc}

\usepackage{ esint }

\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools,esint,color,hyperref,cleveref,subcaption}

\usepackage{amsmath,amsfonts,amssymb,amsthm,color}
\usepackage{mathrsfs}

\usepackage[utf8]{inputenc}
\usepackage{graphicx}

\usepackage{marginnote}

\usepackage[inline]{asymptote}

\usepackage{fancybox,color}

\newcommand{\optf}{f^\star}


\usepackage{todonotes}

\newcommand{\bra}[1]{\langle #1 |}
\newcommand{\ket}[1]{| #1 \rangle}
\newcommand{\braket}[2]{\langle #1 | #2 \rangle}


\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{fact}[theorem]{Fact}
\newtheorem{claim}[theorem]{Claim}
%\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
%\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{Question}[theorem]{Question}
\newtheorem{example}[theorem]{Example}
\newtheorem{conjecture}[theorem]{Conjecture}


\newcommand{\CR}{{\mathcal{R}}}
\newcommand{\R}{{\mathbb R}}
\newcommand{\Q}{{\mathbb Q}}
\newcommand{\N}{{\mathbb N}}
\newcommand{\Z}{{\mathbb Z}}
\newcommand{\C}{{\mathcal C}}
\newcommand{\M}{{\mathcal M}}
\newcommand{\D}{{\mathcal D}}
\newcommand{\E}{{\mathbb E\,}}
\newcommand{\ha}{{\mathcal{H}}}
\newcommand{\CB}{{\mathcal B}}

\newcommand{\newr}[1]{\textcolor{red}{#1}}
\newcommand{\newb}[1]{\textcolor{blue}{#1}}
\newcommand{\new}[1]{\textcolor{black}{#1}}

\title{On classical inequalities for autocorrelations and autoconvolutions
%The inequality Jose and João proposed has extremizers
}
\author{Jaume de Dios Pont}
\address[JDP]{Department of  Mathematics,  University  of  California  Los  Angeles,  Portola Plaza 520, Los  Angeles,
CA 90095, USA}
\email{jdedios@math.ucla.edu}
\author[Jos\'e Madrid]{Jos\'e Madrid}
\address[JM]{Department of  Mathematics,  University  of  California  Los  Angeles,  Portola Plaza 520, Los  Angeles,
CA 90095, USA}
\email{jmadrid@math.ucla.edu}

\date{\today}
\newtheorem{thm}{Theorem}
\newtheorem{prop}{Proposition} 

\keywords{autocorrelation, autoconvolution, optimal constants, extremizers}
\subjclass[2010]{42A05, 42A85, 70H03, 39A12}



\begin{document}
	\maketitle
	
\begin{abstract}
In this paper we study an autocorrelation inequality proposed by Barnard and Steinerberger \cite{BarnardSteinerberger}. The study of these problems is motivated by a classical problem in additive combinatorics. We establish the existence of extremizers to this inequality, for a general class of weights, including Gaussian functions (as studied by the second author and Ramos) and characteristic function (as originally studied by Barnard and Steinerberger). Moreover, via a discretization argument and numerical analysis, we find some almost optimal approximation for the best constant allowed in this inequality. We also discuss some other related problem about autoconvolutions.


 .
\end{abstract}	
	
	
	
\section{Introduction}
\input{intro}


\section{The extremizer exists and is compactly supported}
\input{existence}

\section{Approximation inequalities}
\input{approximation}


\input{computational}

\section{A remark on another autoconvolution problem}
\input{another_problem}





\begin{thebibliography}{99}

\bibitem{BarnardSteinerberger} 
\newblock R.C. Barnard, S. Steinerberger,
\newblock \emph{Three convolution inequalities on the real line with connection to additive combinatorics.}
\newblock Journal of Number Theory, \textbf{207} (2020), 42--55. 

\bibitem{Beckner}
\newblock W. Beckner, 
\newblock \emph{Inequalities in Fourier analysis.} 
\newblock  Annals of Mathematics, \textbf{102} (1975), 159--182. 

\bibitem{BCK}
\newblock  J. Bourgain, L. Clozel and J.P. Kahane, 
\newblock \emph{Principe d’Heisenberg et fonctions positives.}
\newblock Annales de l’institut Fourier, \textbf{60} (2010), n. 4, 1215--1232.

\bibitem{CDGJLM}
 A. Chang, J. de Dios, R. Greenfeld, A. Jamneshan, Z. Li and J. Madrid,
 \newblock \emph{Decoupling for fractal subsets of the parabola.}
\newblock Preprint https://arxiv.org/abs/2012.11458

\bibitem{CJLL}
 E. Carlen, I. Jauslin, E. Lieb and M. Loss,
 \newblock \emph{On the Convolution Inequality $f\geq f*f$.}
\newblock To appear in IMRN https://doi.org/10.1093/imrn/rnaa350.

\bibitem{CRT}
\newblock J. Cilleruelo, I. Ruzsa and C. Trujillo,
\newblock \emph{Upper and lower bounds for finite $Bh[g]$ sequences.}
\newblock Journal of Number Theory, \textbf{97} (2002), n. 1, 26--34.

\bibitem{CRV} 
\newblock J. Cilleruelo, I. Ruzsa and C. Vinuesa,
\newblock \emph{Generalized Sidon sets.}
\newblock Advances in Mathematics, \textbf{225} (2010), n. 5, 2786 --2807.

\bibitem{CS}
\newblock A. Cloninger and S. Steinerberger, 
\newblock \emph{On suprema of autoconvolutions with an application to Sidon sets.}
\newblock Proceedings of the American Mathematical Society \textbf{145} (2017), no. 8, 3191--3200.

\bibitem{CG}
\newblock H. Cohn and F. Gon\c{c}alves,
\newblock \emph{An optimal uncertainty principle in twelve dimensions via modular forms.}
\newblock Inventionnes Mathematicae {\bf 217} (2019), no.~3, 799--831. 

\bibitem{Github}
\newblock  J. de Dios, J. Madrid, 
\newblock \emph{On classical inequalities for autocorrelations and autoconvolutions: accompaining code.}
\newblock Hosted on GitHub at \url{https://github.com/jaumededios/suprema-autocorrelations}.

\bibitem{FKM}
\newblock S. Fish, D. King and S. J. Miller,  
\newblock \emph{Extensions of Autocorrelation Inequalities with Applications to Additive Combinatorics.}
\newblock arXiv preprint \texttt{arXiv:2001.02326}. 

\bibitem{Folland}
\newblock G. Folland, 
\newblock \emph{A course in abstract harmonic analysis.} 
\newblock Chapman and Hall/CRC, 2016. 

%\bibitem{GOeSR2}
%\newblock F. Gon\c{c}alves, D. Oliveira e Silva, and J. P. G. Ramos, 
%\newblock \emph{New sign uncertainty principles.}
%\newblock Preprint, 2020.

%\bibitem{GOeSR1}
%\newblock F. Gon\c{c}alves, D. Oliveira e Silva, and J. P. G. Ramos, 
%\newblock \emph{On regularity and mass concentration phenomena for the sign uncertainty principle.}
%\newblock Preprint, 2020.

\bibitem{GOeSS}
\newblock F. Gon\c{c}alves, D. Oliveira e Silva, and S. Steinerberger,
\newblock \emph{ Hermite polynomials, linear flows on the torus, and an uncertainty principle for roots.} 
\newblock Journal of Mathematical Analysis and Applications {\bf 451} (2017), no.~2, 678--711. 

\bibitem{G}
\newblock B. Green, 
\newblock \emph{The number of squares and $Bh[g]$ sets.}
\newblock Acta Arithimetica, \textbf{100} (2001), 365--390.

\bibitem{HLP}
\newblock G.H. Hardy, J.E. Littlewood, and G. Polya,
\newblock \emph{Inequalities}
\newblock Cambridge University Press, Cambridge (1934).

\bibitem{Ha}
\newblock J. K. Haugland, 
\newblock \emph{Advances in the minimum overlap problem.}
\newblock Journal of Number Theory, \textbf{58} (2001), 71--78.

\bibitem{MOB1}
\newblock G. Martin and K. O'Bryant, 
\newblock \emph{Constructions of generalized Sidon sets.}
\newblock Journal of Combinatorial Theory Series A \textbf{113} (2006), n. 4, 591--607.

\bibitem{MOB2} 
\newblock G. Martin and K. O'Bryant,
\newblock \emph{The symmetric subset problem in continuous Ramsey theory.}
\newblock Experimental Mathematics \textbf{16} (2007), n. 2, 145--166.

\bibitem{MOB3}
\newblock G. Martin, K. O’Bryant,
\newblock \emph{The supremum of autoconvolutions, with applications to additive
number theory.}
\newblock Illinois J. Math. \textbf{53} (2009), no. 1, 219--235.

\bibitem{MR}
\newblock J. Madrid and J. P. G. Ramos, 
\newblock \emph{On optimal autocorrelation inequalities on the real line.}
\newblock Communications on Pure and Applied Analysis (2021), 20 (1) 369-388



\bibitem{MV} 
\newblock  M. Matolcsi and C. Vinuesa, 
\newblock \emph{Improved bounds on the supremum of autoconvolutions.} 
\newblock Journal of Mathematical Analysis and Applications \textbf{372} (2010), no. 2, 439--447.

\bibitem{Y}
\newblock  G. Yu, 
\newblock \emph{An upper bound for $B2[g]$ sets.}
\newblock Journal of Number Theory \textbf{122} (2007), n. 1, 211--220. 


\end{thebibliography}


\end{document}