
\section[Proof of Theorem 5.16]{Proof of Theorem~\ref{thm-ellipse}}
\label{appendix-skew-red}

This appendix contains a proof of Theorem~\ref{thm-ellipse}. We start
by reformulating the problem in more convenient terms. Recall that the
notion of uprightness introduced in Section~\ref{ssec-uprightness} was
defined for an arbitrary bounded convex subset of $\R^2$. If the set
in question is an ellipse, we can expand the definition of uprightness
into an explicit expression. Recall from Definition~\ref{def-ellipses}
that an ellipse centered at $p$ can be written as $E=\s{u\in\R^2\such
  (u-p)^\dagger D(u-p)\leq 1}$, where $D$ is a positive definite matrix
whose entries are, e.g., as follows:
\[
  D = 
  \left[ 
    \begin{array}{cc} 
      a & b  \\
      b & d
    \end{array} 
  \right].
\]
We can compute the area of $E$ and the area of its 
bounding box using $D$. Indeed, we have 
$\area(E)= \pi/\sqrt{\Det(D)}$ and 
$\area (\BBox(E))=4\sqrt{ad} /\Det(D)$. 
Substituting these in Definition~\ref{def-uprightness} 
yields the desired expression for uprightness:
\begin{equation}
\label{eqn:upright}
  \up(E) = \frac{\area(E)}
                {\area(\BBox(E))}
         = \frac{\pi}{4} 
             \sqrt{
               \frac{\Det(D)}{ad} 
             }.
    {\hspace{1.5cm}
    \begin{tikzpicture}[baseline=0, xscale=1.6, yscale=1]
      \draw[fill=yellow!20] (-1,-1) -- (1,-1) -- (1,1) -- (-1,1) -- cycle;
      \draw[fill=red!50, rotate=45, yscale=0.4] (0,0) circle (1.3131);
      \draw (0,0) node {\small $E$};
      \draw (-1,.75) node[right] {\small $\BBox(E)$};
    \end{tikzpicture}
    }
\end{equation}
It follows that the uprightness of $E$ is invariant under 
translation and scalar multiplication.

Recall that $\lambda=\sqrt{2} +1$. The matrix $D$ corresponding to 
an ellipse $E$ has determinant 1 if and only if it can be written 
in the form
\begin{equation}
  D =
  \left[ 
    \begin{array} {cc} 
      e{\lambda^{-z}} & b \\
      b & e{\lambda^{z}}
    \end{array} 
  \right]
\end{equation}
for some $b,e,z\in\R$ with $e>0$ and $e^2=b^2+1$.  In this case, the
definition of uprightness {\eqref{eqn:upright}} simplifies to
\begin{equation}\label{eqn-upe}
  \up(E) = \frac{\pi}{4e^2} = \frac{\pi}{4\sqrt{b^2+1}}.
\end{equation}
Equivalently, if $\up(E)=M$, then
\begin{equation}\label{eqn-b2}
  b^2 = \frac{\pi^2}{16M^2} - 1.
\end{equation}

Since Theorem~\ref{thm-ellipse} deals with pairs of ellipses, it is
convenient to introduce the following terminology for discussing pairs
of matrices.

\begin{definition}
A \emph{state} is a pair of real symmetric positive 
definite matrices of determinant 1. Given a state 
$\state$ with
\begin{equation}
\label{eqn:state}
  D =
  \left[ 
    \begin{array}{cc} 
      e{\lambda^{-z}} & b \\
      b & e{\lambda^{z}}
    \end{array} 
  \right]
  ~~~~
  \Delta =
  \left[ 
    \begin{array}{cc} 
      \varepsilon{\lambda^{-\zeta}} & \beta \\
      \beta & \varepsilon{\lambda^{\zeta}}
    \end{array} 
  \right]
\end{equation}
we define its \emph{skew} as 
$\sk\state=b^2+\beta^2$ and its
\emph{bias} as $\bias \state 
=\zeta-z$.
\end{definition}

Note that the skew of a state is small if and only if both $b^2$ and
$\beta^2$ are small, which happens, by {\eqref{eqn-upe}}, if and only
if the ellipses corresponding to $D$ and $\Delta$ both have large
uprightness. So our strategy for increasing the uprightness will be to
reduce the skew. In what follows, we use $\state$ to denote an
arbitrary state and always assume that the entries of $D$ and $\Delta$
are given as in \eqref{eqn:state}. For future reference, we record
here another useful property of states.

\begin{remark}\label{rem-be}
If $\state$ is a state with $b\geq 0$,
then $-be\leq -b^2$. Indeed:
\[
e^2=b^2+1 ~\imp~ e^2\geq b^2 ~\imp~ e\geq b ~\imp~ -be\leq -b^2.
\]
Similarly, if $b\leq 0$, then $be\leq -b^2$. Analogous 
inequalities also hold for $\beta$ and $\varepsilon$.
\end{remark}

The action of a grid operator on an ellipse can be adapted to states
in a natural way, provided that the operator is special.

\begin{definition}
\label{definition_action}
The action of special grid operators on states is defined as
follows. Here, $\G\da$ denotes the transpose of $\G$, and $\G\bul$ is
defined by applying $(-)\bul$ separately to each matrix entry, as in
Remark~\ref{gridoperatorcomposition}.
\[
\state\cdot \G = (\G^\dagger D\G,\G^{\bullet\dagger} \Delta \G^\bullet).
\]
\end{definition}

\begin{lemma}\label{lem-ellipse-action}
  Let $(D,\Delta)$ be a state, and let $A$ and $B$ be the ellipses
  centered at the origin that are defined by $D$ and $\Delta$,
  respectively.  Then the ellipses $\G(A)$ and $\G\bul(B)$ are defined
  by the matrices $D'$ and $\Delta'$, where
  \[ (D',\Delta') = (D,\Delta) \cdot \G\inv
  \]
\end{lemma}

\begin{proof}
  We have 
  \[ \begin{array}{rcl}
    \G(A) &=& \s{\G(u)\in\R^2 \mid u\da D u \leq 1} \\
    &=& \s{v\in\R^2 \mid (\G\inv v)\da D(\G\inv v) \leq 1} \\
    &=& \s{v\in\R^2 \mid v\da (\G\inv)\da D \G\inv v \leq 1},
  \end{array}
  \] 
  so the ellipse $\G(A)$ is defined by the positive operator
  $D'=(\G\inv)\da D \G\inv$. The proof for $\G\bul(B)$ is similar.
\end{proof}

The main ingredient in the proof of 
Theorem~\ref{thm-ellipse} is the following Step Lemma.

\begin{lemma}[Step Lemma]
\label{Step}
For any state $\state$, if $\sk\state \geq \formula{\P}$, 
then there exists a special grid operator $\G$ such that 
$\sk (\state \cdot \G)\leq \formula{\Qstep} ~\sk\state$. 
Moreover, $\G$ can be computed using a constant number of arithmetic
operations.
\end{lemma}

Before proving the Step Lemma, we show how it can be 
used to derive Theorem~\ref{thm-ellipse}, whose 
statement we reproduce here. 

\begin{un-theorem}
  Suppose $A,B\subseteq \R^2$ are ellipses. Then there exists a grid
  operator $\G$ such that $\G(A)$ and $\G^\bullet (B)$ are
  $1/\formula{\oneoverM}$-upright. Moreover, if $A$ and $B$ are
  $M$-upright, then $\G$ can be efficiently computed in $O(\log(1/M))$
  arithmetic operations.
\end{un-theorem}

\begin{proof}
  Let $D$ and $\Delta$ be the matrices defining $A$ and $B$
  respectively, in the sense of Definition~\ref{def-ellipses}. Since
  uprightness is invariant under translations and scaling, we may
  without loss of generality assume that both ellipses are centered at
  the origin, and that $\det D = \det \Delta = 1$.

  The pair $\state$ is a state. By applying Lemma~\ref{Step}
  repeatedly, we get grid operators $\G_1,\ldots,\G_n$ such that:
  \begin{equation}\label{eqn-skew-reduction}
    \sk(\state\cdot \G_1\ldots \G_n) \leq \formula{\P}.
  \end{equation}
  Now let $(D',\Delta')=\state\cdot \G_1\ldots \G_n$ and set
  $\G=(\G_1\cdots \G_n)^{-1}$. By Lemma~\ref{lem-ellipse-action}, the
  ellipses $\G(A)$ and $\G^{\bullet}(B)$ are defined by the matrices
  $D'$ and $\Delta'$, respectively.  Let $b$ and $\beta$ be the
  anti-diagonal entries of the matrices $D'$ and $\Delta'$,
  respectively.  We have:
  \[
  b^2+\beta^2  = \sk (D',\Delta')=
  \sk(\state\cdot \G\inv) =
  \sk (\state\cdot \G_1\ldots \G_n)\leq \formula{\P},
  \]
  hence $b^2\leq \formula{\P}$ and $\beta^2\leq\formula{\P}$.
  Using {\eqref{eqn-upe}}, we get
  \[
  \up(\G(A)) 
  = \frac{\pi}{4\sqrt{b^2+1}} 
  \geq \frac{\pi}{4\sqrt{\formula{\P+1}}}
  \geq 1/\formula{\oneoverM} 
  \quad\mbox{and}\quad
  \up(\G\bul(B)) 
  = \frac{\pi}{4\sqrt{\beta^2+1}} 
  \geq \frac{\pi}{4\sqrt{\formula{\P+1}}} 
  \geq 1/\formula{\oneoverM},
  \]
  as desired. 
  % Check!!!
  \assert{\P <= (pi^2 / (16 * \M^2))-1}%

  To bound the number of operations, note that each application of
  $\G_j$ reduces the skew by at least $\formula{\roundone{100-100*\Qstep}}$ percent.
  Therefore, the number $n$ in {\eqref{eqn-skew-reduction}} satisfies
  $n\leq \log_{\formula{\Qstep}}({\formula{\P}}/{\sk\state}) =
  O(\log(\sk\state))$.  Using {\eqref{eqn-b2}}, we have
  \[ \log (\sk\state) = \log (b^2+\beta^2)
  \leq \log ((\frac{\pi^2}{16M^2} - 1) + (\frac{\pi^2}{16M^2} - 1))
  = O(\log(1/M)).
  \]
  It follows that the computation of $\G$ requires
  $O(\log(1/M))$ applications of the Step Lemma, each of
  which requires a constant number of arithmetic operations, proving
  the final claim of the theorem. 
\end{proof}

The remainder of this appendix is devoted to proving the 
Step Lemma. To each state, we associate the pair $(z, 
\zeta)$. The proof of the Step Lemma is essentially 
a case distinction on the location of the pair $(z, 
\zeta)$ in the plane. We find 
coverings of the plane with the property that if the 
point $(z, \zeta)$ belongs to some region 
$\mathcal{O}$ of our covering, then it is easy to compute a 
special grid operator $\G$ such that $\sk(\state\cdot \G)\leq 
\formula{\Qstep}~\sk\state$. The relevant grid operators are 
given in Figure~\ref{list_operators}.
%
\begin{figure}
\[
R= \frac{1}{\sqrt{2}}\left[ \begin{array} {cc} 
1 & -1  \\
1 & 1
\end{array} \right]
~~
A = \left[ \begin{array} {cc} 
1 & -2  \\
0 & 1
\end{array} \right]
~~ 
B = \left[ \begin{array} {cc} 
1 & \sqrt{2}  \\
0 & 1
\end{array} \right]
\]
\[
K = \frac{1}{\sqrt{2}}\left[ \begin{array} {cc} 
-\lambda^{-1} & -1  \\
\lambda & 1
\end{array} \right]
~~
X= \left[ \begin{array} {cc} 
0 & 1  \\
1 & 0
\end{array} \right]
~~ 
Z= \left[ \begin{array} {cc} 
1 & 0  \\
0 & -1
\end{array} \right]
\]
\caption{List of useful grid operators.}
\label{list_operators}
\rule{\textwidth}{0.1mm}
\end{figure}
%
Each one of the next 5 subsections is dedicated to a
particular region of the plane. We prove the Step Lemma 
in Section~\ref{ssect-step}.

\subsection{The Shift Lemma}
\label{ssect-shift}

In this section, we consider states $\state$ such that 
$|\bias\state|>1$. Any such state can be ``shifted" to a 
state $(D', \Delta')$ of equal skew but with 
$|\bias(D', \Delta')|\leq 1$.

\begin{definition}
The \emph{shift operators} $\sigma$ and $\tau$ are defined by:
\[
\sigma =
\sqrt{\lambda^{-1}}\left[\begin{array}{cc}
\lambda & 0 \\
0 & 1
\end{array} \right],
\tau =
\sqrt{\lambda^{-1}}
\left[\begin{array}{cc}
1 & 0 \\
0 & -\lambda
\end{array} \right]
\]
\end{definition}

Even though $\sigma$ and $\tau$ are not grid operators, we can use 
them to define an operation on states called a \emph{shift by $k$}. 
By abuse of notation, we write this operation as an action.

\begin{definition}
Given a state $\state$ and $k\in\Z$, the \emph{k-shift of $\state$} is defined as:
\[
\state \cdot \shift^k = (\sigma^k D \sigma^k, \tau^k \Delta \tau^k).
\]
\end{definition}

The notation $\state \cdot \shift^k$ is justified by the following lemma.

\begin{lemma}
\label{propshift}
The shift of a state is a state and moreover:
\[
\sk(\state\cdot\shift^k)= \sk\state
~\mbox{ and }~
\bias(\state\cdot\shift^k)=\bias\state +2k
\]
\end{lemma}

\begin{proof}
Compute $\state\cdot\shift^k$:
\[
\begin{array}{rcl}
\state\cdot\shift^k & = & (\sigma^k D \sigma^k, \tau^k \Delta \tau^k) \\
& = & (\sigma^k\left[ \begin{array} {cc} 
e{\lambda^{-z}} & b \\
b & e{\lambda^{z}}
\end{array} \right] \sigma^k,
\tau^k\left[ \begin{array} {cc} 
\varepsilon{\lambda^{-\zeta}} & \beta \\
\beta & \varepsilon{\lambda^{\zeta}}
\end{array} \right]\tau^k) \\
& = & (\left[ \begin{array} {cc} 
e\lambda^{-z+k} & b \\
b & e\lambda^{z-k}
\end{array} \right],
\left[ \begin{array} {cc} 
\varepsilon\lambda^{-\zeta-k} & (-1)^k\beta \\
(-1)^k\beta & \varepsilon\lambda^{\zeta+k}
\end{array} \right])
\end{array}
\]
The resulting matrices are clearly symmetric and positive definite. Moreover, since $\sigma^k$ and $\tau^k$ have determinant $\pm 1$, both $\sigma^k D\sigma^k$ and $\tau^k \Delta\tau^k$ have determinant 1. Finally:
\begin{itemize}
  \item $\sk(\state\cdot\shift^k)= b^2+((-1)^k\beta)^2=b^2+\beta^2= \sk\state$ and
  \item $\bias(\state\cdot\shift^k)=(\zeta+k)-(z-k)=\bias\state +2k$,
\end{itemize}
which completes the proof.
\end{proof}

For every special grid operator $\G$, there is a special grid operator $\G'$ whose action on a state corresponds to shifting the state by $k$, applying $\G$ and then shifting the state by $-k$.

\begin{lemma}
\label{conjugationbysigma1}
If $\G$ is a special grid operator and $k\in\Z$, then $\G'=\sigma^k \G \sigma^k$ is a special grid operator and moreover $\G'^\bullet = (-\tau)^k \G^\bullet \tau^k$.
\end{lemma}

\begin{proof}
It suffices to show this for $k=1$. Suppose $\G=\left[ \begin{array} {cc} 
w & x \\
y & z
\end{array} \right]$ is a special grid operator and note that:
\[
\G'=
\sigma \G \sigma=
\left[ \begin{array} {cc} 
\lambda w & x \\
y & \lambda^{-1}z
\end{array} \right]
=
\left[ \begin{array} {cc} 
\lambda^{-1} & 0 \\
0 & \lambda^{-1}
\end{array} \right]
\left[ \begin{array} {cc} 
\lambda & 0 \\
0 & 1
\end{array} \right]
\G
\left[ \begin{array} {cc} 
\lambda  & 0 \\
0 & 1
\end{array} \right].
\]
Since all the factors in the above product are grid operators, the result is also a grid operator. Moreover $\Det(\sigma\G\sigma)=\Det(\G)=1$ so that $\sigma\G\sigma$ is special. Finally:
\[
\G'^\bullet=(\sigma \G \sigma)^\bullet=
\left[ \begin{array} {cc} 
\lambda^\bullet w^\bullet & x^\bullet \\
y^\bullet & (\lambda^{-1})^\bullet z^\bullet
\end{array} \right]
=
\left[ \begin{array} {cc} 
-\lambda^{-1} w^\bullet & x^\bullet \\
y^\bullet & -\lambda z^\bullet
\end{array} \right]
=
-\tau \G^\bullet \tau.
\]
\end{proof}

\begin{lemma}
\label{conjugationbysigma2}
If $\G$ is a grid operator, then:
\[
((\state\cdot \shift^k) \cdot \G )\cdot \shift^k = \state \cdot (\sigma^k \G\sigma^k).
\]
\end{lemma}

\begin{proof} Write $\G'=\sigma^k \G\sigma^k$. Simple computation then yields the result:
\[
\begin{array}{rcl}
((\state\cdot \shift^k) \cdot \G )\cdot \shift^k 
& = & ((\sigma^k D \sigma^k,\, \tau^k \Delta \tau^k)\cdot \G )\cdot \shift^k \\
& = & (\G^\dagger\sigma^k D \sigma^k \G,\, \G^{\bullet\dagger} \tau^k \Delta \tau^k \G^\bullet )\cdot \shift^k \\
& = & (\sigma^k\G^\dagger\sigma^k D \sigma^k \G\sigma^k,\, \tau^k \G^{\bullet\dagger} \tau^k \Delta \tau^k \G^\bullet \tau^k ) \\
& = & (\sigma^k\G^\dagger\sigma^k D \sigma^k \G\sigma^k,\, ((-\tau)^k\G^{\bullet\dagger} \tau^k )\Delta ((-\tau)^k \G^\bullet \tau^k) ) \\
& = & (\G'^\dagger D \G',\, \G'^{\bullet\dagger} \Delta {\G'}^\bullet ) \\
& = & \state \cdot \G' \\
& = & \state \cdot (\sigma^k \G\sigma^k).
\end{array}
\]
\end{proof}

Shifts allow us to consider only states $\state$ with 
$\bias\state\in[-1,1]$ in the proof of the Step Lemma.

\begin{lemma}
\label{shift_lemma}
If the Step Lemma holds for all states $\state$ with
$\bias\state\in [-1,1]$, then it holds for all states.
\end{lemma}

\begin{proof}
Let $\state$ be some state with $\sk\state\geq\formula{\P}$. 
Let $x=\bias\state$ and set $k=\floor{\frac{1-x}{2}}$. Then by 
Lemma~\ref{propshift}, we have $\sk(\state\cdot\shift^k)= \sk\state$ and
$\bias(\state\cdot\shift^k)\in[-1,1]$. Then by assumption, 
there exists a special grid operator $\G$ such that 
$\sk((\state\cdot\shift^k)\cdot \G)\leq \formula{\Qstep}~\sk(\state 
\cdot\shift^k)$. Now by Lemma~\ref{conjugationbysigma1} 
we know that $\G'=\sigma^k ~\G~ \sigma^k$ is a special 
grid operator. Moreover, by 
Lemma~\ref{conjugationbysigma2} and \ref{propshift}, 
we have:
\[
\begin{array}{rcl}
\sk(\state\cdot \G') & = & \sk(((\state\cdot\shift^k)\cdot \G)\cdot \shift^k) \\
& = & \sk((\state \cdot\shift^k)\cdot \G) \\
& \leq & \formula{\Qstep}~\sk(\state \cdot\shift^k) \\
& = & \formula{\Qstep}~\sk\state ,
\end{array}
\]
which completes the proof.
\end{proof}

\subsection[The R Lemma]{The $R$ Lemma}
\label{ssect-R}

\begin{definition}
  The \emph{hyperbolic sine in base $\lambda$} is defined as:
\[
  \sinl(x) = \frac{\lambda^x-\lambda^{-x}}{2}.
\]
\end{definition}

\begin{lemma}
\label{lemmaR}
Recall the operator $R$ from Figure~\ref{list_operators}. 
If $\state$ is such that $\sk\state\geq\formula{\P}$, and
such that $\formula{-\r}\leq z\leq\formula{\r}$ and $\formula{-\r}\leq
\zeta\leq\formula{\r}$, then:
\[
  \sk(\state\cdot R) \leq \formula{\Qstep}~\sk\state. 
\]
\end{lemma}

\begin{proof}
Compute the action of $R$ on $\state$:
\[
  \begin{array}{lcl}
    R^\dagger DR 
    & = & \displaystyle\frac{1}{2}
    \left[ \begin{array} {cc} 
      1 & 1  \\
      -1 & 1
    \end{array} \right]
    \left[ \begin{array} {cc} 
      e{\lambda^{-z}} & b \\
      b & e{\lambda^{z}}
    \end{array} \right]
    \left[ \begin{array} {cc} 
      1 & -1  \\
      1 & 1
    \end{array} \right] \\\\[-1.5ex]
    & = &
    \left[ \begin{array} {cc} 
      \ldots & \frac{e({\lambda^{z}}-{\lambda^{-z}})}{2}\\
      \frac{e({\lambda^{z}}-{\lambda^{-z}})}{2} & \ldots
    \end{array} \right] 
    ~ = ~ 
    \left[ \begin{array} {cc} 
      \ldots & e\sinl(z)\\
      e\sinl(z) & \ldots
    \end{array} \right], 
  \\\\[-1.5ex]
    R^{\bullet\dagger} \Delta R^\bullet 
    & = & \displaystyle\frac{1}{2}
    \left[ \begin{array} {cc} 
      -1 & -1  \\
      1 & -1
    \end{array} \right]
    \left[ \begin{array} {cc} 
      \varepsilon{\lambda^{-\zeta}} & \beta \\
      \beta & \varepsilon{\lambda^{\zeta}}
    \end{array} \right]
    \left[ \begin{array} {cc} 
      -1 & 1  \\
      -1 & -1
    \end{array} \right] \\\\[-1.5ex]
    & = & \left[ \begin{array} {cc} 
      \ldots & \frac{\varepsilon({\lambda^{\zeta}}-{\lambda^{-\zeta}})}{2}\\
      \frac{\varepsilon({\lambda^{\zeta}}-{\lambda^{-\zeta}})}{2} & \ldots
    \end{array} \right] 
    ~ = ~ \left[ \begin{array} {cc} 
      \ldots & \varepsilon\sinl(\zeta)\\
      \varepsilon\sinl(\zeta) & \ldots
    \end{array} \right].
  \end{array}
\]
Therefore $\sk(\state\cdot R)= e^2\sinl^2(z)+ 
\varepsilon^2\sinl^2(\zeta)$. But recall that 
$e^2=b^2+1$ and $\varepsilon^2=\beta^2+1$, so that in 
fact:
\[
  \sk(\state\cdot R) = (b^2+1)\sinl^2(z)+
  (\beta^2+1)\sinl^2(\zeta). 
\]
We assumed 
$\formula{-\r}\leq z,\zeta\leq\formula{\r}$
and this implies that 
$\sinl^2(\zeta), \sinl^2(z) \leq 
\sinl^2(\formula{\r})$. Writing 
$y=\sinl^2(\formula{\r})$ for brevity, and using 
the assumption that $\sk\state\geq\formula{\P}$, 
we get:
\[
  \begin{array}{rcl}
  \sk(\state\cdot R) & = &
  (b^2+1)\sinl^2(z)+
  (\beta^2+1)\sinl^2(\zeta)\\
  & \leq & 
  (b^2+1)y+(\beta^2+1)y\\
  & = & (b^2+\beta^2+2)y \\
  & \leq & \sk\state (1+\frac{2}{\formula{\P}})y.
  \end{array}
\]
This completes the proof, since $(1+\frac{2}
{\formula{\P}})y = (1+\frac{2}
{\formula{\P}})\sinl^2(\formula{\r}) 
\approx \formula{(1+(2/\P))*(\sinhl \r)^2} 
\leq \formula{\Qstep}$.
%% Check!!!
\assert{(1+(2/\P))*(\sinhl \r)^2 <= \Qstep}%
%
\end{proof}

\subsection[The K Lemma]{The $K$ Lemma}
\label{ssect-K}

\begin{definition}
The \emph{hyperbolic cosine in base $\lambda$} is 
defined as:
\[
  \cosl (x) = \frac{\lambda^x+\lambda^{-x}}{2}.
\]
\end{definition}

\begin{lemma}
\label{lemmaK}
Recall the operator $K$ from Figure~\ref{list_operators}. 
If $\state$ is such that $\bias\state\in 
[-1,1]$, $\sk\state\geq\formula{\P}$, and such that $b,\beta\geq 0$,
$z\leq\formula{-\a}$, and $\formula{\r}\leq\zeta$, then:
\[
  \sk(\state\cdot K)\leq \formula{\Qstep}~\sk\state.
\]
\end{lemma}

\begin{proof}
Compute the action of $K$ on $\state$:
\begin{eqnarray}
\lefteqn{K^\dagger DK}\nonumber\\
& = & \frac{1}{2}
  \left[ \begin{array}{cc} 
  -\lambda^{-1} & \lambda  \\
  -1 & 1
  \end{array} \right]
  \left[ \begin{array}{cc} 
  e{\lambda^{-z}} & b \\
  b & e{\lambda^{z}}
  \end{array} \right]
  \left[ \begin{array}{cc} 
  -\lambda^{-1} & -1  \\
  \lambda & 1
  \end{array} \right] \nonumber \\[9pt]
& = & \frac{1}{2} 
  \left[ \begin{array}{cc} 
  \ldots & e(\lambda^{z+1}+\lambda^{-z-1})
    -2\sqrt{2}b \\
  e(\lambda^{z+1}+\lambda^{-z-1})-2\sqrt{2}b & \ldots
  \end{array} \right]  \nonumber \\[9pt]
& = & 
  \left[ \begin{array}{cc} 
  \ldots & e\cosl(z+1) -\sqrt{2}b \\
  e\cosl(z+1)  -\sqrt{2}b & \ldots
  \end{array} \right],\nonumber
\end{eqnarray}
\begin{eqnarray} 
\lefteqn{K^{\bullet\dagger} \Delta K^\bullet} 
\nonumber\\
& = & \frac{1}{2}
  \left[ \begin{array}{cc} 
  \lambda & -\lambda^{-1}  \\
  -1 & 1
  \end{array} \right]
  \left[ \begin{array}{cc} 
  \varepsilon{\lambda^{-\zeta}} & \beta \\
  \beta & \varepsilon{\lambda^{\zeta}}
  \end{array} \right]
  \left[ \begin{array}{cc} 
  \lambda & -1  \\
  -\lambda^{-1} & 1
  \end{array} \right] \nonumber \\[9pt]
& = & \frac{1}{2} 
  \left[ \begin{array}{cc} 
  \ldots & -\varepsilon(
    \lambda^{\zeta-1}+\lambda^{-\zeta+1})+2\sqrt{2}\beta \\
  -\varepsilon(\lambda^{\zeta-1}+\lambda^{-\zeta+1})
    +2\sqrt{2}\beta & \ldots
  \end{array} \right]  \nonumber \\[9pt]
& = &
  \left[ \begin{array}{cc} 
  \ldots & \sqrt{2}\beta - \varepsilon \cosl(\zeta-1) \\
  \sqrt{2}\beta - \varepsilon\cosl(\zeta-1)
    & \ldots
  \end{array} \right].\nonumber
\end{eqnarray}

\noindent
Therefore:
\begin{equation}\label{k0}
\sk(\state \cdot K)=(\sqrt{2}b -e\cosl(z+1))^2 
+(\sqrt{2}\beta-\varepsilon\cosl(\zeta-1))^2.
\end{equation}

\noindent
But recall that $e^2=b^2+1$, and from Remark~\ref{rem-be} that $b\geq 0$ 
implies $-be\leq -b^2$, so:
\begin{eqnarray}
\lefteqn{(\sqrt{2}b -e\cosl(z+1))^2} 
\nonumber\\
& = & 2b^2-2\sqrt{2}\,be\cosl(z+1)+ 
e^2\cosl^2(z+1) \nonumber\\
& \leq & 2b^2-2\sqrt{2}\,b^2\cosl(z+1)+ 
(b^2+1)\cosl^2(z+1) \nonumber\\
& = & b^2(2-2\sqrt{2}\cosl(z+1)+ 
\cosl^2(z+1))+\cosl^2(z+1) 
\nonumber\\
& = & b^2(\sqrt{2}-\cosl(z+1))^2+
\cosl^2(z+1).\label{k1}
\end{eqnarray}
Reasoning analogously, we also have
\begin{eqnarray}
  (\sqrt{2}\beta-\varepsilon\cosl(\zeta-1))^2
  & \leq &
  \beta^2(\sqrt{2}-\cosl(\zeta-1))^2+\cosl^2(\zeta-1).
\end{eqnarray}

\noindent
By assumption, $\bias\state\in [-1,1]$, thus 
$\zeta\leq z+1$.
This, together 
with the assumptions $\formula{\r}\leq\zeta$ 
and $z\leq\formula{-\a}$,
implies that both $z+1$ and $\zeta-1$ are in the interval
$[\formula{-\rr},\formula{-\aa}]$. On this interval, the function 
$\cosl^2(x)$ assumes its maximum at $x=\pgfmathparse{
  ifthenelse((\coshl {-\rr}) <= (\coshl {-\aa}), -\aa,
  -\rr)}\pgfmathresult$, and the function 
$f(x) = (\sqrt 2 - \cosl(x))^2$ assumes its maximum at 
$x=\assert{-\rr <= 0 && 0 <= -\aa}% 
\pgfmathparse{
  ifthenelse((\kone{-\rr}) <= (\kone{0}) && (\kone{-\aa}) <=
  (\kone{0}), 0,  
  ifthenelse ((\kone {-\rr}) <= (\kone {-\aa}), -\aa, -\rr)}\pgfmathresult$.
Therefore, 
\begin{eqnarray}
b^2(\sqrt{2}-\cosl(z+1))^2  + \cosl^2(z+1)
& \leq &
\pgfmathparse{
  ifthenelse((\kone{-\rr}) <= (\kone{0}) && (\kone{-\aa}) <= (\kone{0}), 0,
  ifthenelse ((\kone {-\rr}) <= (\kone {-\aa}), -\aa, -\rr)
}
b^2(\sqrt{2} -\cosl(\pgfmathresult))^2 
+ 
\pgfmathparse{
  ifthenelse((\coshl {-\rr}) <= (\coshl {-\aa}), -\aa, -\rr)
}
\cosl^2(\pgfmathresult).
\end{eqnarray}
and
\begin{eqnarray}
\beta^2(\sqrt{2}-\cosl(\zeta-1))^2+\cosl^2(\zeta-1)
& \leq & 
\beta^2\pgfmathparse{
  ifthenelse((\kone{-\rr}) <= (\kone{0}) && (\kone{-\aa}) <= (\kone{0}), 0,
  ifthenelse ((\kone {-\rr}) <= (\kone {-\aa}), -\aa, -\rr)
}
(\sqrt{2} -\cosl(\pgfmathresult))^2
+
\pgfmathparse{
  ifthenelse((\coshl {-\rr}) <= (\coshl {-\aa}), -\aa, -\rr)
}
\cosl^2(\pgfmathresult).\label{k3}
\end{eqnarray}

\noindent
Combining (\ref{k0})--(\ref{k3}), together 
with the assumption that $\sk\state\geq\formula{\P}$, yields:
\begin{eqnarray}
\sk(\state \cdot K) & = & 
(\sqrt{2}b -e\cosl(z+1))^2 
+(\sqrt{2}\beta-\varepsilon\cosl(\zeta-1))^2 \nonumber\\
& \leq & 
(b^2+\beta^2)\pgfmathparse{
  ifthenelse((\kone{-\rr}) <= (\kone{0}) && (\kone{-\aa}) <= (\kone{0}), 0,
  ifthenelse ((\kone {-\rr}) <= (\kone {-\aa}), -\aa, -\rr)
}
(\sqrt{2} -\cosl(\pgfmathresult))^2
+
2\pgfmathparse{
  ifthenelse((\coshl {-\rr}) <= (\coshl {-\aa}), -\aa, -\rr)
}
\cosl^2(\pgfmathresult) \nonumber\\
& = & 
\sk\state\pgfmathparse{
  ifthenelse((\kone{-\rr}) <= (\kone{0}) && (\kone{-\aa}) <= (\kone{0}), 0,
  ifthenelse ((\kone {-\rr}) <= (\kone {-\aa}), -\aa, -\rr)
}
(\sqrt{2} -\cosl(\pgfmathresult))^2
+
2\pgfmathparse{
  ifthenelse((\coshl {-\rr}) <= (\coshl {-\aa}), -\aa, -\rr)
}
\cosl^2(\pgfmathresult) \nonumber\\
& \leq & 
\sk\state(\pgfmathparse{
  ifthenelse((\kone{-\rr}) <= (\kone{0}) && (\kone{-\aa}) <= (\kone{0}), 0,
  ifthenelse ((\kone {-\rr}) <= (\kone {-\aa}), -\aa, -\rr)
}
(\sqrt{2} -\cosl(\pgfmathresult))^2
+
\frac{2}{\formula{\P}}\pgfmathparse{
  ifthenelse((\coshl {-\rr}) <= (\coshl {-\aa}), -\aa, -\rr)
}
\cosl^2(\pgfmathresult)) \nonumber
\end{eqnarray}
This completes the proof since 
$\pgfmathparse{
  ifthenelse((\kone{-\rr}) <= (\kone{0}) && (\kone{-\aa}) <= (\kone{0}), 0,
  ifthenelse ((\kone {-\rr}) <= (\kone {-\aa}), -\aa, -\rr)
}
(\sqrt{2} -\cosl(\pgfmathresult))^2
+
\frac{2}{\formula{\P}}\pgfmathparse{
  ifthenelse((\coshl {-\rr}) <= (\coshl {-\aa}), -\aa, -\rr)
}
\cosl^2(\pgfmathresult)
\approx\formula{max((\kone {-\rr}),(\kone {-\aa}),(\kone{0})) +
(max((\coshl {-\rr})^2, (\coshl {-\aa})^2))*(2/\P)}
\leq \formula{\Qstep}$.
%% Check!!!
\assert{max((\kone {-\rr}),(\kone {-\aa}),(\kone{0})) + 
(max((\coshl {-\rr})^2, (\coshl {-\aa})^2))*(2/\P) 
<= \Qstep}%
%
\end{proof}

\subsection[The A Lemma]{The $A$ Lemma}
\label{ssect-A}

\begin{definition}
  Let $g(x) = (1-2x)^2$. 
\end{definition}

\begin{lemma}
\label{lemmaA}
Recall the operator $A$ from Figure~\ref{list_operators}. 
If $\state $ is such that $\bias\state
\in[-1,1]$, $\sk\state\geq\formula{\P}$, and such that
$b,\beta\geq 0$ and $\formula{-\a}\leq{z},\zeta$,
then there exists $n\in \Z$ 
such that:
\[
  \sk(\state \cdot A^n) 
  \leq \formula{\Qstep}~\sk\state . 
\]
\end{lemma}

\begin{proof}
Let $c=\min\s{{z},{\zeta}}$ and 
$n=\max\s{1, \floor{\frac{\lambda^{c}}{2}}}$.
Compute the action of $A^{n}$ on $\state$:
\[
  \begin{array}{lcl}
    {A^{n}}^\dagger DA^{n} & = & 
    \left[\begin{array} {cc} 
      1 & 0  \\
      -2n & 1
    \end{array}\right]
    \left[ \begin{array} {cc} 
      e{\lambda^{-z}} & b \\
      b & e{\lambda^{z}}
    \end{array}\right]
    \left[ \begin{array} {cc} 
      1 & -2n  \\
      0 & 1
    \end{array}\right] \\[9pt]
    & = & \left[\begin{array} {cc} 
      \ldots & b-2ne{\lambda^{-z}}  \\
      b-2ne{\lambda^{-z}} & \ldots
    \end{array} \right], \\[18pt]
    {A^{n}}^{\bullet\dagger}
      \Delta{A^{n}}^\bullet 
    & = & {A^{n}}^\dagger \Delta A^{n} \\
    & = & \left[ \begin{array} {cc} 
      \ldots & \beta-2n\varepsilon{\lambda^{-\zeta}}  \\
      \beta-2n\varepsilon{\lambda^{-\zeta}} & \ldots
    \end{array} \right].
  \end{array}
\]
Therefore:
\[
\sk(\state \cdot A^{n}) = 
(b-2ne{\lambda^{-z}})^2+(\beta-2n\varepsilon{\lambda^{-\zeta}})^2
\]
But recall that $e^2=b^2+1$ and
$\varepsilon^2=\beta^2+1$, and from Remark~\ref{rem-be} that 
$b,\beta\geq 0$ implies $-be\leq -b^2$ and 
$-\varepsilon\beta\leq -\beta^2$. Using these 
facts, we can expand the above formula as 
follows:
\begin{eqnarray}
\lefteqn{\sk(\state\cdot A^{n})}\nonumber \\
& = & (b-2ne{\lambda^{-z}})^2+(\beta-2n\varepsilon{\lambda^{-\zeta}})^2 
  \nonumber\\
& = & b^2
-4nbe{\lambda^{-z}}
+4n^2e^2{\lambda^{-2z}}
+\beta^2
-4n\beta\varepsilon{\lambda^{-\zeta}}
+4n^2\varepsilon^2{\lambda^{-2\zeta}}
  \nonumber\\
& \leq & b^2
-4nb^2{\lambda^{-z}}
+4n^2(b^2+1){\lambda^{-2z}}
+\beta^2
-4n\beta^2{\lambda^{-\zeta}}
+4n^2(\beta^2+1){\lambda^{-2\zeta}}
  \nonumber\\
& = & 
b^2(
1
-4n{\lambda^{-z}}
+4n^2{\lambda^{-2z}}
)+\beta^2(
1
-4n{\lambda^{-\zeta}}
+4n^2{\lambda^{-2\zeta}}
)+4n^2({\lambda^{-2z}}+{\lambda^{-2\zeta}})
  \nonumber \\
& = & 
b^2(1-2n\lambda^{-z})^2
+\beta^2(1-2n\lambda^{-\zeta})^2
+4n^2({\lambda^{-2z}}+{\lambda^{-2\zeta}})
  \nonumber \\
& = & b^2g(n{\lambda^{-z}})+\beta^2g(n{\lambda^{-\zeta}})+ 
  4n^2({\lambda^{-2z}}+{\lambda^{-2\zeta}}).\nonumber
\end{eqnarray}

\noindent
Writing $y=\max\s{g(n{\lambda^{-z}}), g(n{\lambda^{-\zeta}})}$ for 
brevity, and using the assumption that 
$\sk\state\geq\formula{\P}$ together with 
the fact that $c\leq z,\zeta$, we 
get:
\begin{eqnarray}
  \sk(\state \cdot A^{n}) & \leq & 
  b^2y+\beta^2y+ 8n^2\lambda^{-2c} \nonumber\\
  & = & \sk\state y+ 8n^2\lambda^{-2c} \nonumber\\
  & \leq & \sk\state (y+ 
  \frac{8}{\formula{\P}}n^2\lambda^{-2c}). \nonumber
\end{eqnarray}
To finish the proof, it remains to show that 
$y+ \frac{8}{\formula{\P}}n^2\lambda^{-2c}\leq 
\formula{\Qstep}$.  There are two cases:
\begin{itemize}
  \item If $\floor{\frac{\lambda^{c}}{2}}\geq 1$, 
  then $\frac{\lambda^{c}}{4}\leq n\leq 
  \frac{\lambda^{c}}{2}$. From $n\leq \frac{\lambda^{c}}{2}$, we have
  $2n\lambda^{-c}\leq 1$, and so  
  $\frac{8}{\formula{\P}}n^2\lambda^{-2c} \leq 
  \frac{2}{\formula{\P}}$. Moreover, because 
  $\bias\state\in[-1,1]$, we have 
  $c \leq  z,\zeta \leq c+1$. Hence
  $\frac{1}{4\lambda} = \frac{\lambda^{c}}{4} \lambda^{-c-1}
  \leq n\lambda^{-c-1}
  \leq n{\lambda^{-z}}, n{\lambda^{-\zeta}} 
  \leq n\lambda^{-c}
  \leq 
  \frac{1}{2}$. On the interval $[\frac{1}{4\lambda},
  \frac{1}{2}]$, the function $g(x)$ assumes its maximum at
  $x=\frac{1}{4\lambda}$. This implies that $y\leq 
  g(\frac{1}{4\lambda})$. This completes the 
  present case since we get:
  \[
  y+ \frac{8}{\formula{\P}}n^2\lambda^{-2c}\leq 
  g(\frac{1}{4\lambda})
  +\frac{2}{\formula{\P}} 
  \approx\formula{(1/(2*\l)-1)^2+2/\P}
  \leq \formula{\Qstep}.
  \]
  \assert{(1/(2*\l)-1)^2+2/\P <= \Qstep}%
  \item If $\floor{\frac{\lambda^{c}}{2}}< 1$, 
  then $n=1$ and $\lambda^c< 2$. From 
$\formula{-\a}\leq c$, we have
  %% Check!!!
  \assert{(ln 0.5)/(ln ((sqrt 2)+1)) <= \a}%
  %
  $\frac{8}{\formula{\P}}n^2\lambda^{-2c}\leq
  \frac{8}{\formula{\P}}\lambda^{\formula{2*\a}}$.  
  Moreover, because $\bias\state\in[-1,1]$, we have
  ${\formula{-\a}}\leq c \leq z,\zeta\leq c+1$. With
  $\lambda^c\leq 2$, this implies that $\frac{1}{2\lambda}\leq
  \lambda^{-c-1}\leq \lambda^{-z},\lambda^{-\zeta}
  \leq\lambda^{\formula{\a}}$. Therefore both $\lambda^{-z}$ and
  $\lambda^{-\zeta}$ are in the interval $[\frac{1}{2\lambda},
  \lambda^{\formula{\a}}]$. On this interval, the function $g(x)$
  assumes its maximum at
  \assert{(1-2*(\l^(\a)))^2 < (1-2*(1/(2*\l)))^2}%
  $x=\frac{1}{2\lambda}$, and therefore $y\leq 
  g(\frac{1}{2\lambda})$. This completes the 
  proof since:
  \[
  y+\frac{8}{\formula{\P}}n^2\lambda^{-2c}\leq 
  g(\frac{1}{2\lambda})
  + \frac{8}{\formula{\P}}\lambda^{\formula{2*\a}}
  \approx\formula{(max(((2*(pow(\l,\a))-1)^2),
  (((pow(\l,(-1)))-1)^2)) + ((8*(pow (\l, 2*\a))) / \P))}
  \leq \formula{\Qstep}.
  \]
  %% Check!!!
  \assert{(max(((2*(pow(\l,\a))-1)^2),
  (((pow(\l,(-1)))-1)^2)) + ((8*(pow (\l, 2*\a))) / \P)) 
  <= \Qstep}%
  %
  \end{itemize}
\end{proof}

\subsection[The B Lemma]{The $B$ Lemma}
\label{ssect-B}

\begin{definition}
  Let $h(x)=(1-\sqrt{2}x)^2$.
\end{definition}

\begin{lemma}
\label{lemmaB}
Recall the operator $B$ from Figure~\ref{list_operators}. 
If $\state $ is such that $\bias\state\in 
[-1,1]$, $\sk\state\geq\formula{\P}$, and such that $b\leq 0\leq\beta$ 
and $\formula{-\b}\leq z,\zeta$, then 
there exists $n\in\Z$ such that:
\[
  \sk(\state\cdot B^n) \leq \formula{\Qstep}~\sk\state.
\]
\end{lemma}

\begin{proof}
Let $c=\min\s{z, \zeta}$, 
$n=\max\s{1, \floor{\frac{\lambda^{c}}{\sqrt{2}}}}$ 
and compute the action of $B^n$ on $\state$:
\[
\begin{array}{lcl}
{B^n}^\dagger DB^n & = & \left[ \begin{array} {cc} 
1 & 0  \\
\sqrt{2}n & 1
\end{array} \right]
\left[ \begin{array} {cc} 
e{\lambda^{-z}} & b \\
b & e{\lambda^{z}}
\end{array} \right]
\left[ \begin{array} {cc} 
1 & \sqrt{2}n  \\
0 & 1
\end{array} \right] \\[9pt]
& = & \left[ \begin{array} {cc} 
\ldots & b+\sqrt{2}\,ne{\lambda^{-z}}  \\
b+\sqrt{2}\,ne{\lambda^{-z}} & \ldots
\end{array} \right], \\[18pt]
{B^n}^{\bullet\dagger} D{B^n}^\bullet & = & \left[ \begin{array} {cc} 
1 & 0  \\
-\sqrt{2}n & 1
\end{array} \right]
\left[ \begin{array} {cc} 
\varepsilon{\lambda^{-\zeta}} & \beta \\
\beta & \varepsilon{\lambda^{\zeta}}
\end{array} \right]
\left[ \begin{array} {cc} 
1 & -\sqrt{2}n  \\
0 & 1
\end{array} \right] \\[9pt]
& = & \left[ \begin{array} {cc} 
\ldots & \beta-\sqrt{2}\,n\varepsilon{\lambda^{-\zeta}}  \\
\beta-\sqrt{2}\,n\varepsilon{\lambda^{-\zeta}} & \ldots
\end{array} \right].
\end{array}
\]
Therefore:
\[
\sk(\state\cdot B^n) = (b+\sqrt{2}\,ne{\lambda^{-z}})^2+
(\beta-\sqrt{2}\,n\varepsilon{\lambda^{-\zeta}})^2.
\]
But recall that $e^2=b^2+1$, that 
$\varepsilon^2=\beta^2+1$, and from Remark~\ref{rem-be} that 
$b\leq 0\leq \beta$ implies $be\leq -b^2$ and 
$-\beta\varepsilon\leq -\beta^2$. Using these 
facts, we can expand the above formula as 
follows:
\begin{eqnarray}
\lefteqn{\sk(\state\cdot B^n)}\nonumber \\
& = & (b+\sqrt{2}\,ne{\lambda^{-z}})^2+
(\beta-\sqrt{2}\,n\varepsilon{\lambda^{-\zeta}})^2\nonumber\\
& = & 
b^2
+2\sqrt{2}\,nbe{\lambda^{-z}}
+2n^2e^2{\lambda^{-2z}}
+\beta^2
-2\sqrt{2}\,n\beta\varepsilon{\lambda^{-\zeta}}
+2n^2\varepsilon^2{\lambda^{-2\zeta}}
\nonumber\\
& \leq & 
b^2
-2\sqrt{2}\,nb^2{\lambda^{-z}}
+2n^2(b^2+1){\lambda^{-2z}}
+\beta^2
-2\sqrt{2}\,n\beta^2{\lambda^{-\zeta}}
+2n^2(\beta^2+1){\lambda^{-2\zeta}}
\nonumber\\
& = & b^2(
1
-2\sqrt{2}\,n{\lambda^{-z}}
+2n^2{\lambda^{-2z}}
)+\beta^2(
1
-2\sqrt{2}\,n{\lambda^{-\zeta}}
+2n^2{\lambda^{-2\zeta}}
)+2n^2({\lambda^{-2z}}+{\lambda^{-2\zeta}})\nonumber \\
& = & 
b^2(1-\sqrt{2}\,n\lambda^{-z})^2
+\beta^2(1-\sqrt{2}\,n{\lambda^{-\zeta}})^2
+ 2n^2({\lambda^{-2z}}+{\lambda^{-2\zeta}}).\nonumber \\
& = & b^2h(n{\lambda^{-z}})+\beta^2h(n{\lambda^{-\zeta}})+ 2n^2({\lambda^{-2z}}+{\lambda^{-2\zeta}}).\nonumber
\end{eqnarray}

\noindent
Writing $y=\max\s{h(n{\lambda^{-z}}), h(n{\lambda^{-\zeta}})}$ for
brevity, and using the assumption that $\sk\state\geq\formula{\P}$,
together with the fact that $c\leq z,\zeta$, we get:
\begin{eqnarray}
\sk(\state \cdot B^n) & \leq & 
b^2y+\beta^2y+ 4n^2\lambda^{-2c} \nonumber\\
& = & \sk\state y +4n^2\lambda^{-2c}\nonumber\\ 
&\leq & \sk\state (y +\frac{4}{\formula{\P}}n^2\lambda^{-2c}) 
\nonumber.
\end{eqnarray}

\noindent
To finish the proof, it remains to show that 
$y+\frac{4}{\formula{\P}}n^2\lambda^{-2c}\leq \formula{\Qstep}$.  There
are two cases:
\begin{itemize}
  \item If $\floor{\frac{\lambda^{c}}{\sqrt{2}}}\geq 1$, 
  then $\frac{\lambda^{c}}{2\sqrt{2}}\leq n\leq 
  \frac{\lambda^{c}}{\sqrt{2}}$. From $n\leq
  \frac{\lambda^{c}}{\sqrt{2}}$, we have $2n^2\lambda^{-2c}\leq 1$, and
  so
  $\frac{4n^2\lambda^{-2c}}{\formula{\P}}\leq \frac{2}
  {\formula{\P}}$. Moreover, because $\bias\state\in[-1,1]$, we have
  $c\leq z,\zeta\leq c+1$. Hence $\frac{1}{2\sqrt{2}\,\lambda}
  = \frac{\lambda^{c}}{2\sqrt{2}}\lambda^{-c-1}\leq n\lambda^{-c-1}\leq
  n{\lambda^{-z}}, n{\lambda^{-\zeta}} \leq n\lambda^{-c}\leq
  \frac{1}{\sqrt{2}}$. On the interval $[\frac{1}{2\sqrt{2}\,\lambda},
  \frac{1}{\sqrt{2}}]$, the function $h(x)$ assumes its maximum at
  $x=\frac{1}{2\sqrt{2}\,\lambda}$. This implies that $y\leq
  h(\frac{1}{2\sqrt{2}\,\lambda})$. This
  completes the present case since we get:
  \[
  y+ \frac{4}{\formula{\P}}n^2\lambda^{-2c}\leq 
  h(\frac{1}{2\sqrt{2}\,\lambda}) +\frac{2}{\P} 
  \approx \formula{(1 - sqrt(2)*(1/(2*sqrt(2)*\l)))^2 + (2/\P)}
  \leq \formula{\Qstep}.
  \]
  % !!! Check
  \assert{(1 - sqrt(2)*(1/(2*sqrt(2)*\l)))^2 + (2/\P) <= (\Qstep)}%
  %
  \item If $\floor{\frac{\lambda^{c}}{\sqrt{2}}}< 1$, 
  then $n=1$ and $\lambda^c<\sqrt{2}$. 
  From $\formula{-\b}\leq c$, we have  $\frac{4}{\formula{\P}}n^2\lambda^{-2c}
  \leq \frac{4}{\formula{\P}}\lambda^{\formula{2*\b}}$. Moreover, because
  %% Check!!!
  %  \assert{((ln 1/(sqrt 2)) / (ln \l)) <= \b}%
  %
  $\bias\state\in[-1,1]$, we have
  $\formula{-\b}\leq c\leq z,\zeta\leq c+1$.
  With $\lambda^c\leq\sqrt{2}$, this implies that
  $\frac{1}{\sqrt{2}\,\lambda} \leq \lambda^{-c-1}\leq {\lambda^{-z}}, {\lambda^{-\zeta}} \leq 
  \lambda^{\formula{\b}}$. Therefore both $\lambda^{-z}$ and
  $\lambda^{-\zeta}$ are in the interval
  $[\frac{1}{\sqrt{2}\,\lambda},\lambda^{\formula{\b}}]$.
  On this interval, the function $h(x)$ assumes its maximum at
  \assert{(1- sqrt(2)*(1/(sqrt(2)*\l)))^2 <= (1-sqrt(2)*((\l)^(\b)))^2}%
  $x=\lambda^{\formula{\b}}$, and therefore $y\leq
  h(\lambda^{\formula{\b}})$. 
  This completes the proof since:
  \[
  y+\frac{4}{\formula{\P}}n^2\lambda^{-2c}\leq 
  h(\lambda^{\formula{\b}})
  + \frac{4}{\formula{\P}}\lambda^{\formula{2*\b}}
  \approx\formula{((max((1 - sqrt(2)*(1/(sqrt(2)*\l)))^2,
  (1-sqrt(2)*((\l)^(\b)))^2) + (4/\P * (\l^(2*\b))))}
  \leq \formula{\Qstep}.
  \]
  % Check !!!
  \assert{((max((1 - sqrt(2)*(1/(sqrt(2)*\l)))^2,
  (1-sqrt(2)*((\l)^(\b)))^2) + (4/\P * (\l^(2*\b)))) 
  <= (\Qstep)}%
  %
\end{itemize}
\end{proof}

\subsection{Proof of the Step Lemma}
\label{ssect-step}

The proof of the Step Lemma is now basically a case distinction, using
the cases enumerated in Sections~\ref{ssect-shift}--\ref{ssect-B}, as
well as some additional symmetric cases. In particular, the following
remark will allow us to use the grid operators $X$ and $Z$ to reduce
the number of cases to consider.

\begin{remark}
\label{rem-XZ}
The grid operator $Z$ negates the anti-diagonal entries
of a state while the operator $X$ swaps the diagonal entries
of a state. This follows by simple computation:
  \[
    \state\cdot Z = 
    \left(
      \left[
        \begin{array}{cc}
          e{\lambda^{-z}} & -b \\
          -b & e{\lambda^{z}}
        \end{array} 
      \right]
    ,
      \left[
        \begin{array}{cc}
          \varepsilon{\lambda^{-\zeta}} & -\beta \\
          -\beta & \varepsilon{\lambda^{\zeta}}
        \end{array} 
      \right]
    \right),
  ~~~~
    \state\cdot X = 
    \left(
      \left[
        \begin{array}{cc}
          e{\lambda^{z}} & b \\
          b & e{\lambda^{-z}}
        \end{array} 
      \right]
      ,
      \left[
        \begin{array}{cc}
          \varepsilon{\lambda^{\zeta}} & \beta \\
          \beta & \varepsilon{\lambda^{-\zeta}}
        \end{array} 
      \right]
    \right).
  \]            
  Moreover, $\bias (\state\cdot Z) = \bias \state$ and 
  $\bias (\state\cdot X) = -\bias \state$.
\end{remark}



\begin{un-lemma}[Step Lemma]
For any state $\state$, if $\sk\state \geq \formula{\P}$, 
then there exists a special grid operator $\G$ such that 
$\sk (\state\cdot \G)\leq \formula{\Qstep} ~\sk\state$.
Moreover, $\G$ can be computed using a constant number of arithmetic
operations.
\end{un-lemma}

\begin{proof}
Let $\state$ be a state such that $\sk\state \geq 
\formula{\P}$. By Lemma~\ref{shift_lemma} we can assume 
w.l.o.g. that $\bias\state\in[-1,1]$. Moreover, by 
Remark~\ref{rem-XZ}, we can also assume that $\beta \geq 0$ 
and $z+\zeta\geq 0$. Note that the application of the grid operators
$X$ and/or $Z$ in Remark~\ref{rem-XZ} preserves the fact that 
$\bias\state\in[-1,1]$. We now treat in turn the cases 
$b\geq 0$ and $b\leq 0$.
\begin{description}
  \item[Case 1] \label{b_beta_plus} $b \geq 0$. A covering of 
  the strip defined by $z-\zeta\in[-1,1]$ and 
  $z+\zeta\geq 0$ is depicted in Figure~\ref{fig-covers}(a). 
  The $R$ region (in green) and the $A$ region (in red) are 
  defined as the intersection of this space with 
  $\s{(z,\zeta)~|~\formula{-\r}\leq z,\zeta \leq\formula{\r}}$ and 
  $\s{(z,\zeta)~|~z\leq\formula{-\a}\mbox{ and }\formula{\r}\leq \zeta}$
  respectively. The $K$ and $K^\bullet$ regions (both in blue) 
  fill the remaining space.
  % This covering requires -r <= a.
  %% Check!!!
  \assert{-\r <= \a}%
  %      
  \begin{figure}
  \[ (a)~ \mp{0.9}{\begin{tikzpicture}[scale=1.5]
      % Regions K: [-0.3, \infty] x [-\infty, -0.8]
      \fill[fill=blue!25] (-0.3,-0.8) -- (-0.3,-2.4) .. controls (0.3,-2.4) and
      (0.5,-1.5) .. (0.5,-0.8) -- cycle;
      \draw (0.25,-1.2) node {$XK$};
      \begin{scope}[scale=-1]
        \fill[fill=blue!25] (-0.3,-0.8) -- (-0.3,-2.4) .. controls (0.3,-2.4) and
        (0.5,-1.5) .. (0.5,-0.8) -- cycle;
        \draw (0.15,-1.3) node {$K$};
      \end{scope}
      \begin{scope}[cm={0,1,1,0,(0,0)}]
        \fill[fill=blue!25] (-0.3,-0.8) -- (-0.3,-2.4) .. controls (0.3,-2.4) and
        (0.5,-1.5) .. (0.5,-0.8) -- cycle;
        \draw (0.15,-1.3) node {$XK\bul$};
      \end{scope}
      \begin{scope}[cm={0,-1,-1,0,(0,0)}]
        \fill[fill=blue!25] (-0.3,-0.8) -- (-0.3,-2.4) .. controls (0.3,-2.4) and
        (0.5,-1.5) .. (0.5,-0.8) -- cycle;
        \draw (0.2,-1.3) node {$K\bul$};
      \end{scope}
      % 
      % Region A: [-\infty, -0.3] x [-\infty, -0.3]
      \fill[fill=red!25] (-2.4,-0.3) -- (-0.3,-0.3) -- 
      (-0.3,-2.4) .. controls (-2,-2.4) and (-2.4,-2) ..
      (-2.4,-0.3) -- cycle;
      \draw (-2.5,-0.3) -- (-0.3,-0.3) -- (-0.3,-2.5);
      \draw (-1.2, -1.2) node {\small $XA^n$};
      % 
      % Region XA: [-\infty, -0.3] x [-\infty, -0.3]
      \begin{scope}[scale=-1]
        \fill[fill=red!25] (-2.4,-0.3) -- (-0.3,-0.3) -- 
        (-0.3,-2.4) .. controls (-2,-2.4) and (-2.4,-2) ..
        (-2.4,-0.3) -- cycle;
        \draw (-2.5,-0.3) -- (-0.3,-0.3) -- (-0.3,-2.5);
        \draw (-1.2, -1.2) node {\small $A^n$};
      \end{scope}
      % Region R: [-0.8, 0.8] x [-0.8, 0.8]
      \filldraw[fill=green!25] (0.8, 0.8) -- (-0.8, 0.8) -- (-0.8, -0.8) -- (0.8, -0.8) -- cycle;
      \draw (0.4, 0.4) node {\small $R$};
      \begin{scope}
        \path[clip] (-1.5,-2.5) -- (2.5,1.5) -- (1.5,2.5) -- (-2.5,-1.5) -- cycle;
        \begin{scope}[yscale=-1]
          \fill[fill=blue!50] (-0.3,-2.4) -- (-0.3,-0.8) -- 
          (0.3,-0.8) -- (0.3,-2.4) -- cycle;
          \draw ((-0.3,-2.5) -- (-0.3,-0.8) -- (0.3,-0.8) -- (0.3,-2.5);
        \end{scope}
        \begin{scope}[cm={0,-1,-1,0,(0,0)}]
          \fill[fill=blue!50] (-0.3,-2.4) -- (-0.3,-0.8) -- 
          (0.3,-0.8) -- (0.3,-2.4) -- cycle;
          \draw ((-0.3,-2.5) -- (-0.3,-0.8) -- (0.3,-0.8) -- (0.3,-2.5);
        \end{scope}
        % 
        % Region XA: [-\infty, -0.3] x [-\infty, -0.3]
        \begin{scope}[scale=-1]
          \fill[fill=red!50] (-2.4,-0.3) -- (-0.3,-0.3) -- 
          (-0.3,-2.4) .. controls (-2,-2.4) and (-2.4,-2) ..
          (-2.4,-0.3) -- cycle;
          \draw (-2.5,-0.3) -- (-0.3,-0.3) -- (-0.3,-2.5);
          \draw (-1.2, -1.2) node {\small $A^n$};
        \end{scope}
        % Region R: [-0.8, 0.8] x [-0.8, 0.8]
%        \filldraw[fill=green!50] (0.8, 0.8) -- (-0.8, 0.8) -- (-0.8, -0.8) -- (0.8, -0.8) -- cycle;
        \filldraw[fill=green!50] (0.8, 0.8) -- (-0.8, 0.8) -- (0.8, -0.8) -- cycle;
        \draw (0.4, 0.4) node {\small $R$};
      \end{scope}  
      % 
      % Coordinate system
      \draw[->] (-2.5,0) -- (2.5, 0) node[right] {$z$};
      \draw[->] (0,-2.5) -- (0, 2.5) node[above] {$\zeta$};
      \foreach \x in {-2, -1, 1, 2} {
        \draw (\x,0) -- (\x,-0.1);
      }
      \foreach \y in {-2, -1, 1, 2} {
        \draw (0,\y) -- (-0.1,\y);
      }
      \draw (1,-0.1) node[below] {\small $1$};
      \draw (-0.1,1) node[left] {\small $1$};
      % 
      % Lines for shift lemma
      \draw (-1.5,-2.5) -- (2.5,1.5);
      \draw (-2.5,-1.5) -- (1.5,2.5);
      % Line for z+zeta >= 0
      \draw [dashed] (-2, 2) -- (-0.5,0.5);    
      \begin{scope}[scale=-1]
      \draw [dashed] (-2, 2) -- (-0.5,0.5);          
      \end{scope}
    \end{tikzpicture}
  }
  \quad
  (b)~
   \mp{0.9}{\begin{tikzpicture}[scale=1.5]
      % 
      % Region XB: [-\infty, 0.2] x [-\infty, 0.2]
      \fill[fill=red!25] (-2.4,0.2) -- (0.2,0.2) -- 
      (0.2,-2.4) .. controls (-2,-2.4) and (-2.4,-2) ..
      (-2.4,0.2) -- cycle;
      \draw (-2.5,0.2) -- (0.2,0.2) -- (0.2,-2.5);
      \draw (-1.2, -1.2) node {\small $XB^n$};
      % 
      % Region B: [-\infty, 0.2] x [-\infty, 0.2]
      \begin{scope}[scale=-1]
        \fill[fill=red!25] (-2.4,0.2) -- (0.2,0.2) -- 
        (0.2,-2.4) .. controls (-2,-2.4) and (-2.4,-2) ..
        (-2.4,0.2) -- cycle;
        \draw (-2.5,0.2) -- (0.2,0.2) -- (0.2,-2.5);
        \draw (-1.2, -1.2) node {\small $B^n$};
      \end{scope}
      % Region R: [-0.8, 0.8] x [-0.8, 0.8]
      \filldraw[fill=green!25] (0.8, 0.8) -- (-0.8, 0.8) -- (-0.8, -0.8) -- (0.8, -0.8) -- cycle;
      \draw (0.4, 0.4) node {\small $R$};
      \begin{scope}
        \path[clip] (-1.5,-2.5) -- (2.5,1.5) -- (1.5,2.5) -- (-2.5,-1.5) -- cycle;
        % 
        % Region XB: [-\infty, 0.2] x [-\infty, 0.2]
        % 
        % Region B: [-\infty, 0.2] x [-\infty, 0.2]
        \begin{scope}[scale=-1]
          \fill[fill=red!50] (-2.4,0.2) -- (-0.8,0.2) -- (0.2, -0.8) -- 
          (0.2,-2.4) .. controls (-2,-2.4) and (-2.4,-2) ..
          (-2.4,0.2) -- cycle;
          \draw (-2.5,0.2) -- (-0.8,0.2) -- (0.2, -0.8) -- (0.2,-2.5);
          \draw (-1.2, -1.2) node {\small $B^n$};          
        \end{scope}
        % Region R: [-0.8, 0.8] x [-0.8, 0.8]
        \filldraw[fill=green!50] (0.8, 0.8) -- (-0.8, 0.8) -- (0.8, -0.8) -- cycle;
        \draw (0.4, 0.4) node {\small $R$};
      \end{scope}  
      % 
      % Coordinate system
      \draw[->] (-2.5,0) -- (2.5, 0) node[right] {$z$};
      \draw[->] (0,-2.5) -- (0, 2.5) node[above] {$\zeta$};
      \foreach \x in {-2, -1, 1, 2} {
        \draw (\x,0) -- (\x,-0.1);
      }
      \foreach \y in {-2, -1, 1, 2} {
        \draw (0,\y) -- (-0.1,\y);
      }
      \draw (1,-0.1) node[below=1.2ex] {\small $1$};
      \draw (-0.1,1) node[left=1.2ex] {\small $1$};
      % 
      % Lines for shift lemma
      \draw (-1.5,-2.5) -- (2.5,1.5);
      \draw (-2.5,-1.5) -- (1.5,2.5);
      % Line for z+zeta >= 0
      \draw [dashed] (-2, 2) -- (-0.5,0.5);    
      \begin{scope}[scale=-1]
      \draw [dashed] (-2, 2) -- (-0.5,0.5);          
      \end{scope}
    \end{tikzpicture}
  }
  \]
  \caption{(a) A covering of the region $z-\zeta\in [-1,1]$ and 
  $z+\zeta\geq 0$ for the case $b\geq 0$. (b) A covering of the 
  region $z-\zeta\in [-1,1]$ and $z+\zeta\geq 0$ for the case 
  $b\leq 0$.}
  \label{fig-covers}
  \rule{\textwidth}{0.1mm}
  \end{figure}
  We now consider in turn the possible locations of 
  the pair $(z,\zeta)$ in this covering. 
  \begin{enumerate}
    \item If $\formula{-\r}\leq z,\zeta \leq\formula{\r}$, 
    then $\sk(\state\cdot R)\leq \formula{\Qstep}~\sk\state$ by 
    Lemma~\ref{lemmaR}.
    \item \label{partk}If $z\leq\formula{-\a}$ and 
    $\formula{\r}\leq \zeta$, then 
    $\sk(\state\cdot K)\leq \formula{\Qstep}~\sk\state$ by 
    Lemma~\ref{lemmaK}.    
    \item If $\formula{-\a}\leq z,\zeta$, then there exists 
    $n\in\Z$ such that $\sk(\state\cdot A^n)\leq 
    \formula{\Qstep}~\sk\state$ by Lemma~\ref{lemmaA}.
    \item \label{partsk} If $\formula{\r}\leq z$ and 
    $\zeta\leq\formula{-\a}$, then note that 
    $\state\cdot K^\bullet = (\Delta, D)\cdot K$, and 
    therefore by Lemma~\ref{lemmaK}:
    \[
    \sk(\state\cdot K^\bullet) = \sk ((\Delta, D)\cdot K) 
    \leq \formula{\Qstep}~ \sk (\Delta, D)=
    \formula{\Qstep}~ \sk\state. 
    \]    
  \end{enumerate}
  \item[Case 2] $b\leq 0$. As above, we use a covering of 
  the strip defined by $z-\zeta\in[-1,1]$ and $z+\zeta\geq 0$ and 
  consider the possible locations of $(z, \zeta)$ in this 
  space. The relevant covering is depicted in 
  Figure~\ref{fig-covers}(b), where the $R$ region (in green) 
  is defined as above and the $B$ region (in red) is defined 
  as the intersection of the strip with 
  $\s{(z,\zeta)~|~z,\zeta\geq \formula{-\b}}$.
  % This covering requires -b <= r-1.  
  %% Check!!!
  \assert{-\b <= \r -1}%
  %
  \begin{enumerate}
    \item If $\formula{-\r}\leq z,\zeta\leq\formula{\r}$, 
    then $\sk(\state\cdot R)\leq \formula{\Qstep}~\sk\state$ by 
    Lemma~\ref{lemmaR}.
    \item If $z,\zeta\geq \formula{-\b}$ then there exists 
    $n\in\Z$ such that $\sk(\state\cdot B^n)\leq 
    \formula{\Qstep}~\sk\state$ by Lemma~\ref{lemmaB}.
  \end{enumerate}
\end{description}
Finally, note that only a constant number of calculations are required
to decide which of the above cases applies. Moreover, each case only
requires a constant number of operations. Specifically, the
computation of $k$ and $\sigma^k$ in Lemma~\ref{shift_lemma}, of $n$
and $A^n$ in Lemma~\ref{lemmaA}, and of $n$ and $B^n$ in
Lemma~\ref{lemmaB} each require just a fixed number of operations, and
each of the remaining cases produces a fixed grid operator.
\end{proof}

