\section{Generalized Double Descent}
\label{sec:general}

Here we present a generic (though informal) way of unifying the three types of
double-descent phenomenon we observe. In particular, we introduce a notion of ``Effective Model Complexity'' along which double-descent occurs.

\pnote{Todo, rewrite this.}

\begin{definition}[Learning Algorithm]
A \emph{learning algorithm $\cA$}
$$\cA: (\cX \x \cY)^* \to M$$
is a function mapping $n$ samples $\{(x_i, y_i)\}$
to a \emph{Model} $M: \cX \to \cY$.
\end{definition}
We will consider learning-algorithms which take as input
$n$ i.i.d. samples $(x_i, y_i) \sim \cD$ from a distribution $\cD$,
and output a model $M$ (for example, by training a neural-network via SGD on the samples).


\begin{definition}[Model Complexity]
For a learning-algorithm $\cA$, distribution $\cD$, loss function $\cL$, and $0 \leq \eps \ll 1$,
define the 
\emph{ModelComplexity}$_{\cD, \cL, \eps}(\cA)$
as the maximum number of samples from $\cD$ for which $\cA$ outputs a model with small train loss.
Formally,
$$
\mathrm{ModelComplexity}_{\cD, \cL, \eps}(\cA)
:=
\argmax_n  \Pr_{\substack{
\{(x_i, y_i)\} \sim \cD^n\\
M \gets \cA(\{(x_i, y_i)\})
}}\left[ \frac{1}{n} \sum_i \cL( M(x_i), y_i )  \leq \eps \right] \geq 1-o(1)
$$
\end{definition}

Throughout, we consider learning-algorithms which output models by 
training a neural network, optimizing the train loss via SGD or variants.
In this case, let $\cA_{NN}[\text{Model-Size, Train-Steps}]$ be the learning algorithm
which trains a neural-network of a given size and architecture for the given number of train steps.

For $n$ training samples,
we consider a learning-algorithm $\cA$ to be ``overparmeterized'' if
$$
\mathrm{ModelComplexity}_{\cD, \cL, \eps}(\cA)
\gg n
$$
and ``underparameterized'' if
$\mathrm{ModelComplexity}_{\cD, \cL, \eps}(\cA) \ll n$,
and ``critically parameterized'' if
$\mathrm{ModelComplexity}_{\cD, \cL, \eps}(\cA) \approx n$.


We can now state our main thesis in terms of the above.
\begin{claim}[Phase Transision, Informal]
Neural-network learning algorithms $\cA_{NN}$ undergo
a \emph{phase transition} in their test performance 
in the ``critically parameterized'' regime, when
the \emph{ModelComplexity} is equal to the size of the train set.
That is, when
$$
\mathrm{ModelComplexity}_{\cD, \cL, \eps}(\cA_{NN})
= n
$$
where $n$ is the number of training examples.
\end{claim}


