
\documentclass{article} % For LaTeX2e
\usepackage[table]{xcolor}
\usepackage{ifthen, xkeyval, tikz, calc}

\usepackage{iclr2020_conference,times}

% Optional math commands from https://github.com/goodfeli/dlbook_notation.

\input{math_commands.tex}
\usepackage{hyperref}
\usepackage{url}
\usepackage{graphicx}
% \usepackage{svg}
\usepackage{subcaption}
\usepackage{float}
\usepackage{style}
\usepackage{tikz}
\usepackage{lipsum}
\usepackage{booktabs}
\usepackage{listings}
\lstset{
  basicstyle=\ttfamily,
  columns=fullflexible,
  frame=single,
  breaklines=true,
  postbreak=\mbox{\textcolor{red}{$\hookrightarrow$}\space},
}
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\checkmark}%
\newcommand{\xmark}{\ding{55}}%


\graphicspath{{./figures/}}
%\graphicspath{{./tmp_figures/}}

% \let\marginpar\oldmarginpar
\setlength{\marginparwidth}{5cm}
% \newcommand{\Bnote}[1]{\textbf{[Boaz: #1]}}
\usepackage{xargs}                      % Use more than one optional parameter in a new commands
\usepackage[colorinlistoftodos,prependcaption, textwidth=3cm]{todonotes}


\newcommandx{\gal}[2][1=]{\todo[linecolor=red, backgroundcolor=red!25,bordercolor=red,#1]{#2}}
\newcommandx{\yamini}[2][1=]{\todo[linecolor=green, backgroundcolor=green!25,bordercolor=green,#1]{#2}}
\newcommandx{\Bnote}[2][1=]{\todo[noprepend, linecolor=orange,backgroundcolor=orange!25,bordercolor=orange,#1]{#2}}

%% Notes off via the below:
% \newcommandx{\gal}[2][1=]{}
% \newcommandx{\yamini}[2][1=]{}
% \newcommandx{\Bnote}[2][1=]{}

\newcommand\iwslt{IWSLT`14}

\newcommand{\AdamTxt}{
Optimized using Adam with learning rate $0.0001$ for 4K epochs, and data-augmentation.
}

\newcommand{\SGDTxt}{Optimized using SGD for 500K steps.}
\newtheorem{hypothesis}{Hypothesis}



\title{
Deep Double Descent:\\
Where Bigger Models and More Data Hurt
}

% Authors must not appear in the submitted version. They should be hidden
% as long as the \iclrfinalcopy macro remains commented out below.
% Non-anonymous submissions will be rejected without review.

\author{Preetum Nakkiran\thanks{Work performed in part while Preetum Nakkiran was interning at OpenAI, with Ilya Sutskever.
We especially thank Mikhail Belkin and Christopher Olah for helpful discussions throughout
this work.
Correspondence Email: \texttt{preetum@cs.harvard.edu}}
\\Harvard University \And
Gal Kaplun\thanks{Equal contribution}\\Harvard University
\And
Yamini Bansal\footnotemark[2]\\Harvard University
\And Tristan Yang \\Harvard University \And
Boaz Barak \\Harvard University
% \thanks{ Use footnote for providing further information
% about author (webpage, alternative address)---\emph{not} for acknowledging
% funding agencies.  Funding acknowledgements go at the end of the paper.} \\
\And
\hspace{-1.9in}Ilya Sutskever\\
\hspace{-1.9in}OpenAI
%\texttt{ilyasu@openai.com}
}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to \LaTeX{} to determine where to break
% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
% puts 3 of 4 authors names on the first line, and the last on the second
% line, try using \AND instead of \And before the third author name.

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\iclrfinalcopy % Uncomment for camera-ready version, but NOT for submission.
\listfiles
\begin{document}


\maketitle
\vspace{-0.4cm}
\input{abstract}
\input{intro}
% \input{old-intro}

\input{setup.tex}

\input{model_dd.tex}

\input{epoch_dd}

\input{sample_dd}

% %\input{general_dd.tex}

\input{discussion.tex}


%\input{rffs}
\subsubsection*{Acknowledgments}
% Use unnumbered third level headings for the acknowledgments. All
% acknowledgments, including those to funding agencies, go at the end of the paper.
We thank Mikhail Belkin for extremely useful discussions in the early stages of this work.
We thank Christopher Olah for suggesting the Model Size $\x$ Epoch visualization, which led to the investigation of epoch-wise double descent,
as well as for useful discussion and feedback.
We also thank Alec Radford, Jacob Steinhardt, and Vaishaal Shankar for helpful discussion and suggestions.
P.N. thanks OpenAI, the Simons Institute, and the Harvard Theory Group for a research environment that enabled this kind of work.

We thank Dimitris Kalimeris, Benjamin L. Edelman, and Sharon Qian,
and Aditya Ramesh for comments on an early draft of this work.

This work supported in part by NSF grant CAREER CCF 1452961, BSF grant 2014389, NSF USICCS proposal 1540428, a Google Research award, a Facebook research award, a Simons Investigator Award,
a Simons Investigator Fellowship,
and NSF Awards
CCF 1715187,
CCF 1565264, CCF 1301976, IIS 1409097,
and CNS 1618026. Y.B. would like to thank the MIT-IBM Watson AI Lab for contributing computational resources for experiments.

\bibliography{bib}
\bibliographystyle{iclr2020_conference}

\appendix
%\input{more_discussion.tex}
\input{appendix}

\end{document}
