\section{Experimental Setup}


We briefly describe the experimental setup here; full details are in Appendix~\ref{sec:appendix_exp_details}
\footnote{The raw data from our experiments
are available at:
\url{https://gitlab.com/harvard-machine-learning/double-descent/tree/master}}.
We consider three families of architectures: ResNets, standard CNNs, and Transformers.
\textbf{ResNets:}
We parameterize a family of ResNet18s (\cite{he2016identity})
by scaling the width (number of filters) of convolutional layers.
Specifically, we use layer widths $[k, 2k, 4k, 8k]$ for varying $k$.
The standard ResNet18 corresponds to $k=64$.
\textbf{Standard CNNs:}
We consider a simple family of 5-layer CNNs, with 4 convolutional layers of widths $[k, 2k, 4k, 8k]$ for varying $k$, and a fully-connected layer. For context, the CNN with width $k=64$, can reach over $90\%$ test accuracy on CIFAR-10 with data-augmentation.
\textbf{Transformers:}
We consider the 6 layer encoder-decoder from \cite{Vaswani},
as implemented by \cite{ott2019fairseq}.
%\footnote{https://github.com/pytorch/fairseq}.
We scale the size of the network by
modifying the embedding dimension $d_{\text{model}}$,
and setting the width of the fully-connected layers proportionally
($d_{\text{ff}} = 4\cdot d_{\text{model}}$).
For ResNets and CNNs, we train with cross-entropy loss, and the following optimizers:
(1) Adam with learning-rate $0.0001$ for 4K epochs; (2) SGD with learning rate $\propto \frac{1}{\sqrt{T}}$ for 500K gradient steps.
We train Transformers for 80K gradient steps, with 10\% label smoothing and no drop-out.

\paragraph{Label Noise.}
In our experiments, label noise of probability $p$ refers to
training on a samples which have the correct label
with probability $(1-p)$, and a uniformly random incorrect label otherwise (label noise is sampled only once and not per epoch).
Figure~\ref{fig:errorvscomplexity} plots test error on the noisy distribution, while the remaining figures plot test error with respect to the clean distribution (the two curves are just linear rescaling of one another).