\vspace{1em}
\begin{minipage}[t]{\textwidth}
\begin{minipage}[t]{0.41\textwidth}\vspace{0pt}
    \centering
\setlength{\tabcolsep}{2pt}
\small
\captionof{table}{\small Window attention has poor performance on long text. The perplexity is restored when we reintroduce the initial four tokens alongside the recent 1020 tokens (4+1020). Substituting the original four initial tokens with linebreak tokens ``\textbackslash n" (4"\textbackslash n"+1020) achieves comparable perplexity restoration.
Cache config x+y denotes adding x initial tokens with y recent tokens. Perplexities are measured on the first book (65K tokens) in the PG19 test set.}
\vspace{-1em}
\label{tab:linebreak}
\begin{tabular}{lc}
\toprule
 Llama-2-13B                                &  PPL ($\downarrow$) \\
\midrule
0 + 1024 (Window)                      & 5158.07                        \\
4 + 1020                    & 5.40                         \\
4"\textbackslash n"+1020 &  5.60                         \\
\bottomrule
\end{tabular}
\end{minipage}
  \hfill
\begin{minipage}[t]{0.57\textwidth}\vspace{0pt}
\centering
\setlength{\tabcolsep}{2pt}
\small
    \captionof{table}{\small Effects of reintroduced initial token numbers on \method. (1) Window attention (0+y) has a drastic increase in perplexity. (2) Introducing one or two initial tokens doesn't fully restore model perplexity, showing that the model doesn't solely use the first token as the attention sink.
    (3) Introducing four initial tokens generally suffices; further additions have diminishing returns. Cache config x+y denotes adding x initial tokens to y recent tokens. Perplexities are evaluated on 400K tokens in the concatenated PG19 test set.
    }
    \vspace{-1em}
    \label{tab:start_size}
    \begin{tabular}{lccccc}
    \toprule
Cache Config  & 0+2048& 1+2047 & 2+2046 & 4+2044 & 8+2040 \\
\midrule
Falcon-7B  & 17.90                               & 12.12                      & 12.12                      & 12.12                      & 12.12                       \\
MPT-7B     & 460.29                              & 14.99                      & 15.00                      & 14.99                      & 14.98                       \\
Pythia-12B & 21.62                               & 11.95                      & 12.09                      & 12.09                      & 12.02                       \\
\midrule
\midrule
Cache Config  & 0+4096 & 1+4095 & 2+4094 & 4+4092 & 8+4088 \\
\midrule
Llama-2-7B & 3359.95                             & 11.88                      & 10.51                      & 9.59                       & 9.54                     \\
\bottomrule
\end{tabular}
  \end{minipage}
\end{minipage}
