\documentclass[../hippo.tex]{subfiles}
\begin{document}

\section{HiPPO-LegS Theoretical Properties}
\label{sec:hippo-theory}

\subsection{Timescale equivariance}

\begin{proof}[Proof of \cref{prop:timescale}]%
    Let $\tilde{f}(t) = f(\alpha t)$.
    Let $c = \proj f$ and $\tilde{c} = \proj \tilde{f}$.
    By the HiPPO equation~\eqref{eq:coefficient} update and the basis
    instantiation for LegS (equation \eqref{eq:scale-legendre-p}),
    \begin{align*}%
        \tilde{c}_n(t) &= \langle \tilde{f}, g_n^{(t)} \rangle_{\mu^{(t)}} \\
        &= \int \tilde{f}(t) (2n+1)^{\frac{1}{2}} P_n\left( 2\frac{x}{t} - 1 \right) \frac{1}{t} \mathbb{I}_{[0, 1]}\left(\frac{x}{t}\right) \dd x
        \\
        &= \int f(\alpha t) (2n+1)^{\frac{1}{2}} P_n\left( 2\frac{x}{t} - 1 \right) \frac{1}{t} \mathbb{I}_{[0, 1]}\left(\frac{x}{t}\right) \dd x
        \\
        &= \int f(\alpha t) (2n+1)^{\frac{1}{2}} P_n\left( 2\frac{x}{\alpha t} - 1 \right) \frac{1}{\alpha t} \mathbb{I}_{[0, 1]}\left(\frac{x}{\alpha t}\right) \dd x
        \\
        &= c_n(\alpha t).
    \end{align*}
    The second-to-last equality uses the change of variables $x \mapsto \frac{x}{\alpha}$.
\end{proof}

\subsection{Speed}





In this section we work out the fast update rules according to the forward
Euler, backward Euler, bilinear, or generalized bilinear transform
discretizations (cf.\ \cref{sec:discretization-full}).
Recall that we must be able to perform matrix-vector multiplication by $I + \delta A$ and $(I - \delta A)^{-1}$
where $\delta$ is some multiple of the step size $\Delta t$ (equation~\eqref{eq:gbt}).

It is easily seen that the LegS update rule involves a matrix $A$ of the following form (\cref{thm:legs}):
$A = D_1 (L + D_0) D_2$, where $L$ is the all $1$ lower triangular matrix and $D_0, D_1, D_2$ are diagonal.
Clearly, $I + \delta A$ is efficient (only requiring $O(N)$ operations),
as it only involves matrix-vector multiplication by diagonals $D_0, D_1, D_2$, or multiplication by $L$ which is the $\mathsf{cumsum}$ operation.

Now we consider multiplication by the inverse $(I+\delta A)^{-1}$ (the minus sign can be absorbed into $\delta$).
Write
\begin{align*}
  (I + \delta D_1 (L+D_0) D_2)^{-1}
  &= \left( D_1 (D_1^{-1} D_2^{-1} + \delta (L+D_0)) D_2 \right)^{-1} \\
  &= \delta^{-1} D_2^{-1} \left( \delta^{-1} D_1^{-1} D_2^{-1} + D_0 + L \right)^{-1} D_1^{-1}
\end{align*}
Since diagonal multiplication is efficient,
the crucial operation is inversion multiplication by a matrix of the form $L + D$.

Consider solving the equation $(L + D) x = y$.
This implies $x_0 + \dots + x_{k-1} = y_k - (1+d_k) x_k$.
The solution is
\begin{align*}
  x_0 &= \frac{y_0}{1+d_0} \\
  x_k &= \frac{y_k - x_0 - \dots - x_{k-1}}{1+d_k} \\
\end{align*}
Define $s_k = x_0 + \dots + x_k$.
Then
\begin{align*}
  s_k = s_{k-1} + x_k = s_{k-1} + \frac{y_k - s_{k-1}}{1 + d_k} = \frac{y_k + d_k s_{k-1}}{1 + d_k} = \frac{d_k}{1+d_k}s_{k-1} + \frac{y_k}{1+d_k}.
\end{align*}

Finally, consider how to calculate a recurrence of the following form efficiently.
\begin{align*}
  x_0 = \beta_0, x_k = \alpha_k x_{k-1} + \beta_{k}.
\end{align*}
This update rule can also be written
\begin{align*}
  \frac{x_k}{\alpha_k\dots\alpha_1} = \frac{x_{k-1}}{\alpha_{k-1}\dots\alpha_1} + \frac{\beta_k}{\alpha_k\dots\alpha_1}.
\end{align*}
Evidently $x$ can be computed in a vectorized way as
\begin{align*}
  x = \mathsf{cumsum}(\beta / \mathsf{cumprod}(\alpha)) \cdot \mathsf{cumprod}(\alpha)
  .
\end{align*}
This is an $O(N)$ computation.




\subsection{Gradient Norms}

We analyze the discrete time case under the Euler discretization (\cref{sec:discretization-full}),
where the HiPPO-LegS recurrent update is equation~\eqref{eq:legs-discrete}, restated here for convenience:
\begin{align*}
  c_{k+1} = \left( 1 - \frac{A}{k} \right) c_k + \frac{1}{k} B f_k.
\end{align*}
These gradient asymptotics hold under other discretizations.

We will show that
\begin{proposition}%
  For any times $k < \ell$, the gradient norm of the HiPPO-LegS operator
  for the output at time $\ell+1$ with respect to input at time $k$ is
  $\left\| \frac{\partial c_{\ell+1}}{\partial f_{k}} \right\| = \Theta\left( 1 / \ell \right)$.
\end{proposition}
\begin{proof}%
  We take $N$ to be a constant.

  Without loss of generality assume $k > 2$, as the gradient change for a single initial step is bounded.
  By unrolling the recurrence~\eqref{eq:legs-discrete}, the dependence of $c_{\ell+1}$ on $c_k$ and $f_k, \dots, f_{\ell}$ can be made explicit:
  \begin{align*}
    c_{\ell+1} &= \left( I - \frac{A}{\ell} \right) \dots \left( I - \frac{A}{k} \right) c_k
    \\
    &\quad + \left( I - \frac{A}{\ell} \right) \dots \left( I - \frac{A}{k+1} \right) \frac{B}{k} f_k
    \\
    &\quad + \left( I - \frac{A}{\ell} \right) \dots \left( I - \frac{A}{k+2} \right) \frac{B}{k+1} f_{k+1}
    \\
    &\quad \quad \vdots
    \\
    &\quad + \left( I - \frac{A}{\ell} \right) \frac{B}{\ell-1} f_{\ell-1}
    \\
    &\quad + \frac{B}{\ell} f_{\ell}
    .
  \end{align*}
  Therefore
  \begin{align*}
    \frac{\partial c_{\ell+1}}{\partial f_k} = \left( I - \frac{A}{\ell} \right) \dots \left( I - \frac{A}{k+1} \right) \frac{B}{k}.
  \end{align*}
  Notice that $A$ has distinct eigenvalues $1, 2, \dots, N$, since those are the elements of its diagonal and $A$ is triangular (\cref{thm:legs}).
  Thus the matrices $I - \frac{A}{\ell}, \dots, I - \frac{A}{k+1}$ are diagonalizable with a common change of basis.
  The gradient then has the form $P D P^{-1} B$ for some invertible matrix
  $P$ and some diagonal matrix $D$.
  Its norm is therefore bounded from below (up to constant) by the smallest
  singular value of $P$ and $\|P^{-1} B\|$, both of which are nonzero constants,
  and the largest diagonal entry of $D$.
  It thus suffices to bound this largest diagonal entry of $D$, which is the
  largest eigenvalue of this product,
  \begin{align*}
    \rho = \left(1 - \frac{1}{\ell}\right) \dots \left(1 - \frac{1}{k+1}\right) \frac{1}{k}.
  \end{align*}
  The problem reduces to showing that $\rho = \Theta(1/l)$.

  We will use the following facts about the function $\log \left( 1 - \frac{1}{x} \right)$.
  First, it is an increasing function, so
  \begin{align*}
    \log \left( 1 - \frac{1}{x} \right) \ge \int_{x-1}^x \log \left( 1 - \frac{1}{\lambda} \right) \dd \lambda.
  \end{align*}
  Second, its antiderivative is
  \begin{align*}
    \int \log \left( 1 - \frac{1}{x} \right) = \int \log(x-1) - \log(x) = (x-1)\log(x-1) - x\log(x) = x\log\left( 1-\frac{1}{x} \right) - \log(x-1)
    .
  \end{align*}
  
  Therefore, we have
  \begin{align*}
    \log \left(1 - \frac{1}{\ell}\right) \dots \left(1 - \frac{1}{k+1}\right)
    &= \sum_{i=k+1}^\ell \log (1 - \frac{1}{i})
    \\
    &\ge \sum_{i=k+1}^\ell \int_{i-1}^i \log \left(1 - \frac{1}{x}\right) \dd x
    \\
    &= \int_{k}^\ell \log \left(1 - \frac{1}{x}\right) \dd x
    \\
    &= \left[ (x-1)\log(x-1) - x\log(x) \right] \big|_{k}^{\ell}
    \\
    &= \ell \log \left( 1-\frac{1}{\ell} \right) - \log(\ell-1)
    \\& - \left( k \log \left( 1-\frac{1}{k} \right) - \log(k-1) \right)
    .
  \end{align*}
  Finally, note that $x \log \left( 1 - \frac{1}{x} \right)$ is an increasing function, and bounded from above since it is negative, so it is $\Theta(1)$ (this can also be seen from its Taylor expansion).
  Thus we have
  \begin{align*}
    \log \rho \ge \Theta(1) - \log(\ell-1) + \log(k-1) - \log(k),
  \end{align*}
  Furthermore, all inequalities are asymptotically tight, so that $\rho = \Theta(1 / \ell)$ as desired.
  
\end{proof}




\subsection{Function Approximation Error}

\begin{proof}[Proof of \cref{prop:approximation_error}]%
  Fix a time $t$.
  HiPPO-LegS uses the measure $\omega(t, x) = \frac{1}{t} \mathbb{I}_{[0, t]}$ and the
  polynomial basis $p_n(t, x) = (2n+1)^{\frac{1}{2}} P_n \left( \frac{2x}{t} - 1 \right)$.
  Let $c_n(t) = \langle f_{\leq t}, p_n^{(t)} \rangle_{\mu^{(t)}}$ for $n = 0, 1, \dots$.
  Then the projection $g^{(t)}$ is obtained by linear combinations of the basis
  functions, with $c_n(t)$ as coefficients:
  \begin{equation*}
    g^{(t)}= \sum_{n=0}^{N - 1} c_n(t) p_n^{(t)}.
  \end{equation*}
  Since $p_n^{(t)}$ forms an orthonormal basis of the Hilbert space defined by
  the inner product $\langle \cdot, \cdot \rangle_{\mu^{(t)}}$~\citep{chihara}, by Parseval's
  identity,
  \begin{equation*}
    \norm{f_{\leq t} - g^{(t)}}^2_{\mu^{(t)}} = \sum_{n=N}^{\infty} c_n^2(t).
  \end{equation*}

  To bound the error $\norm{f_{\leq t} - g^{(t)}}_{\mu^{(t)}}$, it suffices to
  bound the sum of the squares of the high-order coefficients $c_n(t)$ for
  $n = N, N+1, \dots$.
  We will bound each coefficient by integration by parts.

  We first simplify the expression for $c_n(t)$.
  For any $n \geq 1$, we have
  \begin{align*}
    c_n(t)
    &= \langle f_{\leq t}, p_n^{(t)} \rangle_{\mu^{(t)}} \\
    &= \frac{1}{t} (2n+1)^{\frac{1}{2}} \int_{0}^{t} f(x) P_n\left(\frac{2x}{t} - 1\right) \d x \\
    &= \frac{(2n+1)^{\frac{1}{2}}}{2} \int_{-1}^{1} f \left( \frac{1+x}{2} t \right) P_n(x) \d x  & \text{(change of variable $x \to \frac{1+x}{2} t$)}.
  \end{align*}

  As $P_n(x) = \frac{1}{2n+1} \frac{d}{dx} (P_{n+1}(x) - P_{n-1}(x))$ (cf.\
  \cref{sec:legendre-properties}), integration by parts yields:
  \begin{align*}
    c_n(t)
    &= \frac{(2n+1)^{\frac{1}{2}}}{2} \left. \left[ f \left( \frac{1+x}{2} t \right) \frac{1}{2n+1} (P_{n+1}(x) - P_{n-1}(x)) \right] \right|_{-1}^1 \\
    &\quad - \frac{(2n+1)^{\frac{1}{2}}}{2} \int_{-1}^{1} \frac{t}{2} f' \left( \frac{1+x}{2} t \right) \frac{1}{2n+1} (P_{n+1}(x) - P_{n-1}(x)) \d x.
  \end{align*}
  Notice that the boundary term is zero, since
  $P_{n+1}(1) = P_{n-1}(1) = 1$ and $P_{n+1}(-1) = P_{n-1}(-1) = \pm 1$ (either
  both 1 or both $-1$ depending on whether $n$ is odd or even).
  Hence:
  \begin{equation*}
    c_n(t) = -\frac{1}{4} \cdot \frac{1}{(2n+1)^{\frac{1}{2}}} \cdot t \int_{-1}^{1} f' \left( \frac{1+x}{2} t \right) (P_{n+1}(x) - P_{n-1}(x)) \d x.
  \end{equation*}

  Now suppose that $f$ is $L$-Lipschitz, which implies that $\abs{f'} \leq L$.
  Then
  \begin{align*}
    c_n^2(t)
    &\leq t^2 L^2 \frac{1}{16} \cdot \frac{1}{2n+1} \left[ \int_{-1}^{1} \abs{P_{n+1}(x) - P_{n-1}(x)} \d x \right]^2 \\
    &\leq t^2 L^2 \frac{1}{16} \cdot \frac{1}{2n+1} \cdot 2 \int_{-1}^{1} (P_{n+1}(x) - P_{n-1}(x))^2 \d x && \text{(Cauchy--Schwarz)} \\
    &= t^2 L^2 \frac{1}{8} \frac{1}{2n+1} \left[ \int_{-1}^{1} P_{n+1}^2(x) \d x + \int_{-1}^{1} P_{n-1}^2(x) \d x \right] && \text{($P_{n+1}$ and $P_{n-1}$ are orthogonal)} \\
    &= t^2 L^2 \frac{1}{8} \frac{1}{2n+1} \left[ \frac{2}{2n+3} + \frac{2}{2n-1} \right] \\
    &= O(1) t^2 L^2 \frac{1}{n^2}.
  \end{align*}
  Summing for all $n \geq N$ yields:
  \begin{align*}
    \norm{f_{\leq t} - g^{(t)}}^2_{\mu^{(t)}}
    &= \sum_{n=N}^{\infty} c_n^2(t) = O(1) t^2 L^2 \sum_{n=N}^{\infty} \frac{1}{n^2} = O(1) t^2 L^2 \frac{1}{N}.
  \end{align*}
  We then obtain that $\norm{f_{\leq t} - g^{(t)}}_{\mu^{(t)}} = O(tL/\sqrt{N})$ as claimed.

  Now supposed that $f$ has $k$ derivatives and the $k$-th derivative is bounded.
  The argument is similar to the one above where we integrate by parts $k$
  times.
  We sketch this argument here.

  Take $k$ to be a constant, and let $n \geq k$.
  Applying integration by parts $k$ times, noting that all the boundary
  terms are zero, gives:
  \begin{equation*}
    c_n(t) = O(1) (2n+1)^{\frac{1}{2}} t^k \int_{-1}^{1} f^{(k)} \left( \frac{1+x}{2} t \right) q_k(x) \d x,
  \end{equation*}
  where $q_k(x)$ is a polynomial such that $\frac{\d^k}{\d x^k} q_k(x) = P_n(x)$.
  Then since $f^{(k)}$ is bounded,
  $\abs{c_n(t)} = O(1) (2n+1)^{\frac{1}{2}} \int_{-1}^{1} \abs{q_k(x)} \d x$, and so
  \begin{equation*}
    c_n^2(t) = O(1) t^{2k} (2n+1) \left[ \int_{-1}^{1} \abs{q_k(x)} \d x \right]^2 = O(1) t^{2k} (2n+1) \int_{-1}^{1} q_k^2(x) \d x \quad \text{(Cauchy--Schwarz)}.
  \end{equation*}
  It remains to bound $\int_{-1}^{1} q_k^2(x) \d x$.
  Using the fact that
  $\frac{\d}{\d x} P_n(x) = \frac{1}{2n+1} (P_{n+1}(x) - P_{n-1}(x))$ repeatedly, we have:
  \begin{align*}
    q_1 &= \frac{1}{2n+1} (P_{n+1} - P_{n-1}) = \frac{1}{n+O(1)} \cdot \frac{1}{2} (P_{n+1} - P_{n-1}) \\
    q_2 &= \frac{1}{(n+O(1))^2} \frac{1}{2^2} (P_{n+2} - P_n - P_n + P_{n-2}) = \frac{1}{(n+O(1))^2} \frac{1}{2^2} (P_{n+2} - 2P_n + P_{n-2}) \\
    q_3 &= \frac{1}{(n+O(1))^3} \frac{1}{2^3} (P_{n+3} - P_{n+1} - 2P_{n+1} + 2P_{n-1} + P_{n-1} - P_{n-3}) = \frac{1}{(n+O(1))^3} \frac{1}{2^3} (P_{n+3} - 3P_{n+1} + 3P_{n-1} - P_{n-3}) \\
    \dots
  \end{align*}
  In general, when we expand out $\int_{-1}^{1} q_k^2(x) \d x$, since the $P_m$'s
  are orthogonal, we get $k+1$ terms of the form
  $\frac{1}{(n+O(1))^{2k}} \frac{1}{2^{2k}} \binom{k}{l}^2 \int_{-1}^{1} P_m^2(x) \d x$
  for $k$ different values of $m$ in the range $[n-k, n+k]$, and $l$ goes from 0 to $k$.
  For each $m$, $\int_{-1}^{1} P_m^2(x) \d x = \frac{1}{n+O(1)}$, and
  $\sum_{l=0}^{k} \binom{k}{l}^2 = \binom{2k}{k}$.
  Summing up all $k+1$ terms yields
  \begin{equation*}
    \int_{-1}^{1} q_k^2(x) \d x = \frac{1}{(n+O(1))^{2k+1}} \frac{1}{2^k} \binom{2k}{k}.
  \end{equation*}
  By Stirling's approximation, $\binom{2k}{k} = O(1) 4^k$, so
  $\int_{-1}^{1} q_k^2(x) \d x = \frac{O(1) 2^k}{(n+O(1))^{2k+1}}$.
  Noting that $k$ is a constant, plugging this into the bound for $c_n^2(t)$:
  \begin{equation*}
    c_n^2(t) = O(1) t^{2k} (2n+1) \frac{O(1)2^k}{(n+O(1))^{2k+1}} = O(1) t^{2k} \frac{1}{n^{2k}}.
  \end{equation*}
  Summing for all $n \geq N$ yields:
  \begin{align*}
    \norm{f_{\leq t} - g^{(t)}}^2_{\mu^{(t)}}
    &= \sum_{n=N}^{\infty} c_n^2(t) = O(1) t^{2k} \sum_{n=N}^{\infty} \frac{1}{n^{2k}} = O(1) t^{2k} \frac{1}{N^{2k-1}}.
  \end{align*}
  We then obtain that $\norm{f_{\leq t} - g^{(t)}}_{\mu^{(t)}} = O(t^k N^{-k+1/2})$ as claimed.
\end{proof}

\textbf{Remark.} The approximation error of Legendre polynomials
reduces to how fast the Legendre coefficients decay, subjected to the smoothness
assumption of the input function.
This result is analogous to the classical result in Fourier
analysis, where the $n$-th Fourier coefficients decay as $O(n^{-k})$ if the input
function has order-$k$ bounded derivatives~\citep{korner1989fourier}.
That result is also proved by integration by parts.

\end{document}


