\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2020

% ready for submission
%\usepackage[nonatbib]{neurips_2020}

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
%\usepackage[preprint]{neurips_2020}

% to compile a camera-ready version, add the [final] option, e.g.:
\usepackage[nonatbib,final]{neurips_2020}

% to avoid loading the natbib package, add option nonatbib:
%\usepackage[nonatbib]{neurips_2020}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage[dvipsnames]{xcolor}
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false, citecolor=ForestGreen]{hyperref}     % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\usepackage{graphicx}
\usepackage{comment}
\usepackage{amsmath,amssymb} % define this before the line numbering.
\usepackage{xspace}
% \usepackage[font=footnotesize,labelfont=bf]{caption} 
\usepackage{caption}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{soul}
\usepackage{xspace}
\usepackage[subrefformat=parens]{subfig}
\usepackage{cleveref}
\usepackage{enumitem}
\usepackage[square,numbers,sort&compress]{natbib}

\usepackage{bbding}
\usepackage{pifont}
\usepackage{wasysym}
\usepackage{amssymb}
\usepackage{float}
\usepackage{dblfloatfix}
\usepackage{footnote}
\makesavenoteenv{tabular}
\makesavenoteenv{table}
\makesavenoteenv{figure}
%\usepackage{authblk}
\usepackage{xspace}
\usepackage{multirow}


\newcommand{\aj}[1]{{\color{red}[\textbf{Armand}:#1]}}
\newcommand{\mc}[1]{{\color{green}[\textbf{Mathilde}:#1]}}
\newcommand{\pb}[1]{{\color{orange}[\textbf{Piotr}:#1]}}
\newcommand{\pg}[1]{{\color{cyan}[\textbf{Priya}:#1]}}
\newcommand{\im}[1]{{\color{blue}[\textbf{Ishan}:#1]}}
\newcommand{\jm}[1]{{\color{magenta}[\textbf{Julien}:#1]}}

% formatting macros
\newcommand{\todo}[1]{{\color{red}\textbf{#1}}}

% \DeclareRobustCommand\onedot{\futurelet\@let@token\@onedot}
\def\onedot{.}

\def\eg{\emph{e.g}\onedot} \def\Eg{\emph{E.g}\onedot}
\def\ie{\emph{i.e}\onedot} \def\Ie{\emph{I.e}\onedot}
\def\cf{\emph{c.f}\onedot} \def\Cf{\emph{C.f}\onedot}
\def\etc{\emph{etc}\onedot} \def\vs{\emph{vs}\onedot}
\def\wrt{w.r.t\onedot} \def\dof{d.o.f\onedot}
\def\etal{\emph{et al}\onedot}
\crefname{section}{\S}{\S\S}
\crefname{subsection}{\S}{\S\S}
\crefformat{table}{Table~#2#1#3}
\crefformat{figure}{Fig.~#2#1#3}
\crefformat{equation}{Eqn~#2#1#3}


% method macros
\def\OURS{SwAV\xspace}
\def\OURSFULL{Swapping Assignments between Views\xspace}
%\newcommand{\appendixName}{supplementary material}
\newcommand{\appendixName}{Appendix}


% arch macros
\newcommand{\resnet}{ResNet\xspace}
\newcommand{\resnext}{ResNeXt\xspace}
\newcommand{\resnetfiftyLong}{ResNet-50\xspace}
\newcommand{\resnetfifty}{R50}
\newcommand{\resnetfiftyVtwo}{R50v2}
\newcommand{\resnethundred}{R101}
\newcommand{\resnetonefifty}{R152}
\newcommand{\resnextfifty}{RX-50}
\newcommand{\resnexthundred}{RX101}
\newcommand{\convone}{\texttt{conv1}}
\newcommand{\resfour}{\texttt{res4}}
\newcommand{\resfive}{\texttt{res5}}

% dataset macros
\newcommand{\ImNetDataset}{ImageNet\xspace}
\newcommand{\ImNet}{ImageNet\xspace}
\newcommand{\ImNetNine}{ImageNet-9k\xspace}
\newcommand{\ImNetFull}{ImageNet-22k\xspace}
\newcommand{\YFCCFull}{YFCC-100M\xspace}
\newcommand{\YFCCone}{YFCC-1M\xspace}
\newcommand{\YFCC}{YFCC\xspace}
\newcommand{\iNat}{iNat18\xspace}
\newcommand{\VOCseven}{VOC07\xspace}
\newcommand{\VOCseventwelve}{VOC07+12\xspace}
\newcommand{\Places}{Places205\xspace}

\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\com}[1]{{\color{CadetBlue}\texttt{#1}}}

\title{
Unsupervised Learning of Visual Features \\
by Contrasting Cluster Assignments
}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the
% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
% authors names on the first line, and the last on the second line, try using
% \AND instead of \And before the third author name.

\author{Mathilde Caron$^{1,2}$ \And Ishan Misra$^2$ \And \hspace{0.2cm} Julien Mairal$^1$ \hspace{-0.2cm} \AND \hspace{0.5cm} Priya Goyal$^2$ \And \hspace*{0.3cm} Piotr Bojanowski$^2$ \And Armand Joulin$^2$ \AND \normalfont $^1$ Inria$^*$ \And \normalfont $^2$ Facebook AI Research}

\begin{document}

\maketitle

\begin{abstract}
Unsupervised image representations have significantly reduced the gap with supervised pretraining, notably with the recent achievements of contrastive learning methods.
These contrastive methods typically work online and rely on a large number of explicit pairwise feature comparisons, which is computationally challenging.
In this paper, we propose an online algorithm, \OURS, that takes advantage of contrastive methods without requiring to compute pairwise comparisons.
Specifically, our method simultaneously clusters the data while enforcing consistency between cluster assignments produced for different augmentations (or ``views'') of the same image, instead of comparing features directly as in contrastive learning.
Simply put, we use a ``swapped'' prediction mechanism where we predict the code of a view from the representation of another view.
Our method can be trained with large and small batches and can scale to unlimited amounts of data.
Compared to previous contrastive methods, our method is more memory efficient since it does not require a large memory bank or a special momentum network.
In addition, we also propose a new data augmentation strategy, \texttt{multi-crop}, that uses a mix of views with different resolutions in place of two full-resolution views, without increasing the memory or compute requirements.
We validate our findings by achieving $75.3\%$ top-1 accuracy on ImageNet with ResNet-50, as well as surpassing supervised pretraining on all the considered transfer tasks.

\end{abstract}

\newcommand\blfootnote[1]{%
  \begingroup
  \renewcommand\thefootnote{}\footnotetext{#1}%
  \addtocounter{footnote}{-1}%
  \endgroup
}
\blfootnote{* Univ. Grenoble Alpes, Inria, CNRS, Grenoble INP, LJK, 38000 Grenoble, France}
\blfootnote{Correspondence to \texttt{mathilde@fb.com}}
\blfootnote{Code: \url{https://github.com/facebookresearch/swav}}

\input{intro.tex}

\input{related.tex}

\input{method.tex}

\input{experiments.tex}

\section{Discussion}
%This work presents an online clustering approach inspired by the recent progress in contrastive learning.
%We show that comparing different views of the same image is not specific to instance classification by proposing a similar principle for clustering.
%We enforce consistency between clustering codes for different augmentations of an image and 
%obtain clustering `codes' for an image using Sinkhorn-Knopp in an online manner.
%Additionally, we propose a \texttt{multi-crop} data augmentation strategy that can be applied to any self-supervised method with consistent gains. 
Self-supervised learning is rapidly progressing compared to supervised learning, even surpassing it on transfer learning, even though the current experimental settings are designed for supervised learning.
In particular, architectures have been designed for supervised tasks, and it is not clear if the same models would emerge from exploring architectures with no supervision.
Several recent works have shown that exploring architectures with search~\cite{liu2020labels} or pruning~\cite{caron2020pruning} is possible without supervision, and we plan to evaluate the ability of our method to guide model explorations.

\paragraph{Acknowledgement.}
We thank Nicolas Carion, Kaiming He, Herve Jegou, Benjamin Lefaudeux, Thomas Lucas, Francisco Massa, Sergey Zagoruyko, and the rest of Thoth and FAIR teams for their help and fruitful discussions.
Julien Mairal was funded by the ERC grant number 714381 (SOLARIS project) and by ANR 3IA MIAI@Grenoble Alpes (ANR-19-P3IA-0003).


\bibliographystyle{splncs04}
\bibliography{egbib}

\newpage
\input{appendix.tex}

\end{document}
