\vspace{-1ex}
\section{Experiment}
\vspace{-0.5ex}
\label{sec:experiment}

\input{table/zeroshot}
\input{table/vqa_zeroshot}

Table 1 provides an overview of the performance of BLIP-2 on various zero-shot vision-language tasks.
Compared to previous state-of-the-art models,
BLIP-2 achieves improved performance while requiring substantially fewer number of trainable parameters during vision-language pre-training.

\vspace{-0.5ex}
\subsection{Instructed Zero-shot Image-to-Text Generation}
\vspace{-0.5ex}
BLIP-2 effectively enables a LLM to understand images while preserving its capability in following text prompts,
which allows us to control image-to-text generation with instructions.
We simply append the text prompt after the visual prompt as input to the LLM.
Figure~\ref{fig:example} shows examples to demonstrate a wide range of zero-shot image-to-text capabilities including visual knowledge reasoning, visual commensense reasoning,
visual conversation,
personalized image-to-text generation, etc.





\textbf{Zero-shot VQA}.
We perform quantitative evaluation on the zero-shot visual question answering task.
For OPT models, we use the prompt ``Question: {\small\{\}} Answer:''.
For FlanT5 models, we use the prompt ``Question: {\small\{\}} Short answer:''.
During generation,
we use beam search with a beam width of 5.
We also set the length-penalty to -1 which encourages shorter answers that align better with human annotation.

As shown in Table~\ref{tbl:vqa_zeroshot}.
BLIP-2 achieves state-of-the-art result on the VQAv2~\cite{VQA2} and GQA~\cite{GQA} datasets.
It outperforms Flamingo80B by 8.7\% on VQAv2, despite having 54x fewer trainable parameters.
On the OK-VQA~\cite{okvqa} dataset,
BLIP-2 comes secondary to Flamingo80B.
We hypothesis that this is because OK-VQA focuses more on open-world knowledge than visual understanding,
and the 70B Chinchilla~\cite{chinchilla} language model from Flamingo80B possesses more knowledge than the 11B FlanT5$_\text{XXL}$.

We make a promising observation from Table~\ref{tbl:vqa_zeroshot}: 
\textbf{a stronger image encoder or a stronger LLM both lead to better performance.}
This observation is supported by several facts:
(1) ViT-g outperforms ViT-L for both OPT and FlanT5.
(2) Within the same LLM family, larger models outperform smaller ones.
(3) FlanT5, an instruction-tuned LLM, outperforms the unsupervised-trained OPT on VQA. 
This observation validates BLIP-2 as a \textbf{generic vision-language pre-training method} that can efficiently harvest the rapid advances in vision and natural language communities.  

% \subsection{Pretrained \name~Prevents Catastrophic Forgetting}
\noindent\textbf{Effect of Vision-Language Representation Learning.}
\vspace{-0.5ex}
\input{figure/qformer_effect.tex}

The first-stage representation learning pre-trains the \name~to learn visual features relevant to the text,
which reduces the burden of the LLM to learn vision-language alignment.
Without the representation learning stage,
\name~relies solely on the vision-to-language generative learning to bridge the modality gap, which is similar to the Perceiver Resampler in Flamingo.
Figure~\ref{fig:qformer_effect} shows the effect of representation learning on generative learning.
Without representation learning,
both types of LLMs give substantially lower performance on zero-shot VQA.
In particular, OPT suffers from catastrophic forgetting where performance drastically degrades as training proceeds.

\input{table/caption}
\input{table/vqa_finetune}


\subsection{Image Captioning}

We finetune BLIP-2 models for the image captioning task,
which asks the model to generate a text description for the image's visual content.
We use the prompt ``a photo of'' as an initial input to the LLM and trains the model to generate the caption with the language modeling loss.
We keep the LLM frozen during finetuning, 
and updates the parameters of the \name~together with the image encoder.
We experiment with ViT-g and various LLMs.
Detailed hyperparameters can be found in the appendix.
We perform finetuning on COCO,
and evaluate on both COCO test set and zero-shot transfer to NoCaps~\cite{nocaps} validation set.

The results are shown in Table~\ref{tbl:caption}.
BLIP-2 achieves state-of-the-art performance with significant improvement on NoCaps over existing methods,
demonstrating strong generalization ability to out-domain images.

\subsection{Visual Question Answering}

Given annotated VQA data,
we finetune the parameters of the \name~and the image encoder while keeping the LLM frozen. 
We finetune with the open-ended answer generation loss,
where the LLM receives \name's output and the question as input,
and is asked to generate the answer.
In order to extract image features that are more relevant to the question,
we additionally condition \name~on the question.
Specifically,
the question tokens are given as input to the \name~and interact with the queries via the self-attention layers,
which can guide the \name's~cross-attention layers to focus on more informative image regions.
% \name's output query embeddings can be conditioned on text input.
% We make use of this property and perform question-guided image feature extraction,
% where the queries interact with the question tokens via self-attention layers.
% This enables the queries to focus on regions in the image that are more relevant to the question,
% thereby feeding the LLM with more informative image features to answering the question.
% Table~\ref{tbl:vqa_ablation} shows that question-guided image feature extraction leads to a noticeable improvement on the VQA performance.

% In order to keep pre-training consistent with finetuning, 
% we perform second-stage pre-training using the language-guided GemFormer,
% where the prefix text is used as input to the GemFormer.
% We also mix VQA data into the second-stage pre-training.

Following BLIP,
our VQA data includes the training and validation splits from VQAv2, as well as training samples from Visual Genome.
Table~\ref{tbl:vqa_finetune} demonstrates the state-of-the-art results of BLIP-2 among open-ended generation models.


% Compare OPT and T5.


\subsection{Image-Text Retrieval}
\input{table/retrieval}
Since image-text retrieval does not involve language generation,
we directly finetune the first-stage-pretrained model w/o LLM.
Specifically,
we finetune the image encoder together with \name~on COCO using the same objectives (\ie~ITC, ITM, and ITG) as pre-training.
We then evaluate the model for both image-to-text retrieval and text-to-image retrieval on COCO and Flickr30K~\cite{flickr} datasets.
During inference,
we follow~\citet{ALBEF,blip} which first select $k=128$ candidates based on the image-text feature similarity, followed by a re-ranking based on pairwise ITM scores.
We experiment with both ViT-L and ViT-g as the image encoder.
Detailed hyperparameters can be found in the appendix.

The results are shown in Table~\ref{tbl:retrieval}.
BLIP-2 achieves state-of-the-art performance with significant improvement over existing methods on zero-shot image-text retrieval.
% On COCO,
% BLIP-2 w/ ViT-g achieves similar performance as BEIT-3, despite having 0.7B fewer parameters during finetuning and 10x fewer trainable parameters during pre-training.


\input{table/retrieval_ablation.tex}
The ITC and ITM losses are essential for image-text retrieval as they directly learn image-text similarity.
In Table~\ref{tbl:retrieval_ablation},
we show that the ITG (image-grounded text generation) loss is also beneficial for image-text retrieval.
This result supports our intuition in designing the representation learning objectives: the ITG loss enforces the queries to extract visual features most relevant to the text,
thus improving vision-language alignment. 






% \subsection{Vision-language Tasks}

% \subsection{Zero-shot VQA}

% \subsection{Fine-tuning for Visual Reasoning}

% \subsection{Multi-task Unified Model}

% \subsection{Visualization}