\section{Limitation}
\label{sec:limitation}
\vspace{-0.5ex}

Recent LLMs can perform in-context learning given few-shot examples.
However,
our experiments with BLIP-2 do not observe an improved VQA performance when providing the LLM with in-context VQA examples.
We attribute the lack of in-context learning capability to our pre-training dataset,
which only contains a single image-text pair per sample.
The LLMs cannot learn from it the correlation among multiple image-text pairs in a single sequence.
The same observation is also reported in the Flamingo paper,
which uses a close-sourced interleaved image and text dataset (M3W) with multiple image-text pairs per sequence.
We aim to create a similar dataset in future work.


BLIP-2's image-to-text generation could have unsatisfactory results due to various reasons including inaccurate knowledge from the LLM,
activating the incorrect reasoning path,
or not having up-to-date information about new image content (see Figure~\ref{fig:example_limitation}).
Furthermore,
due to the use of frozen models,
BLIP-2 inherits the risks of LLMs,
such as outputting offensive language, propagating social bias, or leaking private information.
Remediation approaches include using instructions to guide model's generation or training on a filtered dataset with harmful content removed.
% In the short term,
% designing better instructions to guide model's generation can be an effective remedy.

% Since BLIP-2 is an efficient method that can flexibly leverage new pre-trained models and datasets,
% we will continue to build stronger vision-language models with the help of 

\vspace{-0.5ex}
\section{Conclusion}
\vspace{-0.5ex}
\label{sec:conclusion}
We propose BLIP-2,
a generic and compute-efficient method for vision-language pre-training that leverages frozen pre-trained image encoders and LLMs.
BLIP-2 achieves state-of-the-art performance on various vision-language tasks while having a small amount of trainable parameters during pre-training.
BLIP-2 also demonstrates emerging capabilities in zero-shot instructed image-to-text generation.
We consider BLIP-2 as an important step towards building a multimodal conversational AI agent.