\documentclass[acmtog,authorversion,nonacm]{acmart}

\usepackage{booktabs} % For formal tables

\citestyle{acmauthoryear}
\setcitestyle{square}


\usepackage[ruled]{algorithm2e} % For algorithms
\renewcommand{\algorithmcfname}{ALGORITHM}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}
\usepackage{tikz}
\acmJournal{TOG}

\usepackage{tabularx}
\newcolumntype{C}[1]{>{\centering\arraybackslash}m{#1}}

\acmSubmissionID{463}

\setcopyright{acmlicensed}



\gdef\cropInsets{0}

\graphicspath{{./Figures/}{./figs/result/}}

\input{macros}
\input{scene_names}

\definecolor{hgblue}{RGB}{138,200,224}
\definecolor{hgred}{RGB}{245,138,143}

\definecolor{gold}{RGB}{221, 196, 65}
\definecolor{silver}{RGB}{215, 215, 215}
\definecolor{bronze}{RGB}{126, 66, 5}

\begin{document}

\title{Optimizing Camera Matrices through Rigid-Body Physics}

\author{Thomas M\"uller}
\affiliation{%
  \institution{NVIDIA}
   \city{Z\"urich}
   \country{Switzerland}
  }
\email{tmueller@nvidia.com}

\renewcommand\shortauthors{M\"uller}

\begin{CCSXML}
<ccs2012>
   <concept>
       <concept_id>10010147.10010169.10010170.10010174</concept_id>
       <concept_desc>Computing methodologies~Massively parallel algorithms</concept_desc>
       <concept_significance>300</concept_significance>
       </concept>
   <concept>
       <concept_id>10010147.10010169.10010170.10010173</concept_id>
       <concept_desc>Computing methodologies~Vector / streaming algorithms</concept_desc>
       <concept_significance>300</concept_significance>
       </concept>
   <concept>
       <concept_id>10010147.10010257.10010293.10010294</concept_id>
       <concept_desc>Computing methodologies~Neural networks</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
 </ccs2012>
\end{CCSXML}

\ccsdesc[300]{Computing methodologies~Massively parallel algorithms}
\ccsdesc[300]{Computing methodologies~Vector / streaming algorithms}
\ccsdesc[500]{Computing methodologies~Neural networks}

\keywords{Image Synthesis, Neural Networks, Encodings, Hashing, GPUs, Parallel Computation, Function Approximation. }


\begin{abstract}
bla bla
\end{abstract}

\maketitle

\newpage
\section{Introduction} \label{Sec:Intro}

Recent developments in inverse rendering have not only increased the demand for better scene reconstruction given camera parameters, but also for finding camera parameters given a specified scene (or a joint optimization of both).

However, automatic differentiation onto a camera matric yields malformed rotational components, which prior work has addressed with matrix-logarithm-space parameterizations and screw transforms.

We propose an alternative parameterization---rotational impulses, coupled with a rotational Adam optimizer---that is equivalent to log-space rotation generators while offering efficient, stable, computation through standard vector operations (dot and cross products) and being simple to understand.

\section{Problem Description} \label{Sec:ProblemDescription}

In this document, we define camera matrices as $\cam \in \R^{3\times4}$
\begin{align}
  \cam := \bigl[ \, \camRot \,\,\, \camPos \, \bigl] \,,
\end{align}
consisting of a rotation matrix $\camRot \in \R^{3\times3}$ representing the camera's orientation, and a translation vector $\camPos \in \R^3$ representing its position.

The camera matrix determines the viewpoint of the (inverse) rendering problem in question.
Namely, primary rays with origin $\origin$ and direction $\originDir$ in canonical camera space have the form
\begin{align}
  \pos = \camRot \cdot ( \origin + t \, \originDir ) + \camPos \,.
\end{align}

By a differentiable rendering algorithm (such as NeRF~\citep{mildenhall2020nerf}), we are given gradients of the loss $\Loss$ with respect to positions from primary ray equation, i.e.\ $\partial{\Loss} / \partial{\pos} \in \R^3$ and our goal is translate these gradients into meaningful updates to the camera matrix $\cam$.


\section{Method} \label{Sec:Method}

Updating the camera position $\camPos$ is easy: its effect on the loss is computed trivially as ${\partial{\Loss} / \partial{\camPos} = \partial{\Loss} / \partial{\pos}}$, permitting the use of gradient-based optimizers such as Adam~\citep{KingmaB14}.
In the following, we will therefore focus solely on the question of how to update the camera rotation $\camRot$.

\paragraph{Na{\"i}ve gradient descent.}
The obvious procedure of computing the ${3\!\times\!3}$ gradient matrix ${\partial{\Loss} / \partial{\camRot}}$ and applying Adam immediately proves problematic: the resulting matrix is no longer valid rotation.
Gradient descent is unaware of the fact, that we do not want it to utilize just 3 out of the available 9 degrees of freedom of the linear transform.

\paragraph{Log-space rotations.}
We can enforce this by exploiting the form of log-space rotation matrices. bla bla, prior work did this already but rely on autodiff, which leads to numeric difficulty due to typical implementations of matrix logarithms and exponentials.

\paragraph{Impulse vectors.}
The important realization is that the 3 values of a log-space rotation, when interpreted as a vector, are the \emph{axis of rotation}, with the vector's norm being the angle of rotation.
In other words: the vector corresponds to the rotational \emph{impulse} induced by $\partial{\Loss} / \partial{\pos}$ and can be computed by a single cross product.

Averaging many log-space gradients from a batch thus amounts to averaging their rotational impulses, which has a very intuitive interpretation: if the camera was a rigid body, then the resulting rotational impulse would be overall effect of each training sample's gradient pushing on it.

\paragraph{Rotational Adam.}
It seems tempting to directly apply Adam to impulse vectors and it turns out that this is physically quite meaningful.
bla bla physical angular momentum, bla bla

Lastly, we must be careful with how we represent $\partial{\camRot}$: it must not actually be in terms of an angle axis! (Arguments about commutativity)


\begin{acks}
bla bla
\end{acks}


\bibliographystyle{ACM-Reference-Format}
\bibliography{egbib}


\end{document}
