% DO NOT WRITE HERE, this content has been moved to the models_experiments.tex file 
\section{ALBERT}
In this section, we present a set of ALBERT models and give an overall comparison of these models to the original BERT models. 
We will have detailed analysis of how each component contribute to the differences and the reasons behind the choices of these models in section \ref{sec:experiments}.

\subsection{Model architectures}
BERT uses the popular transformer encoder architecture \citep{vaswani2017attention}, which we will not review in detail. Following the annotation of BERTs,
we denote the number transformer encoder layers as $L$, the hidden size as $H$, the number of self-attention heads as $A$,
and the vocab embedding size as $E$. We also set the feed-forward/filter size to be $4H$. We primary report results on the models listed in Table \ref{table:model_architecture}. 
The major differences between BERT and ALBERT are smaller vocabulary embedding size and cross-layer parameter sharing. Because of these two techniques, ALBERTs have much smaller parameter size compared to BERT with similar architecture. For example, ALBERT-large has about 18x less parameters compared to BERT-large. For BERTs, besides base and large models as reported by \cite{devlin2018bert}, we are able to train an xlarge model with a hidden size 2x as large as the BERT-large model. However, this BERT-xlarge has about 1.3 billions parameters, which make it difficult to train as it requires a Cloud-TPU machine with 1024 chips that is hard to get. On the contrary, ALBERTs have much smaller parameter size hence memory footprint, so we are able to scale to a model with 4x larger hidden size than BERT-large while having smaller parameter size. 




\begin{table}
\small
\centering
\begin{tabular}{l  l | c c c c c c}
 \multicolumn{2}{c}{Model}  &Parameters	&Layers	&Hidden	& Heads	&Embedding &Parameter-sharing \\
\hline
\multirow{3}{*}{BERT}    &base	        &110M    &12 & 768  & 12 & 768 & False\\
    &large	        &340M	 &24 & 1024 & 16 & 1024 & False\\
    &xlarge	&1370M	 &24 & 2048 & 32 & 2048 & False\\
    \hline
\multirow{4}{*}{ALBERT}    &base	        &12M     &12 & 768  & 12 & 128 & True \\
    &large	        &18M	 &24 & 1024 & 16 & 128 & True\\
    &xlarge	&59M	 &24 & 2048 & 32 & 128 & True\\
    &xxlarge	&233M	 &12 & 4096 & 64 & 128 & True\\
\end{tabular}
\caption{The architectures of the primary models reported in this paper.}
\label{table:model_architecture}
\end{table}


\subsection{Overall Performance Comparison}
\label{overall-performance}

\begin{table}
\small
\centering
\begin{tabular}{ l l | c  c c c c c c}
 \multicolumn{2}{c}{Model}            	&SQuAD1.1	&SQuAD2.0	&MNLI-m	&SST-2	&RACE	&Avg & Speedup\\
\hline
\multirow{3}{*}{BERT} &base	          &90.5/83.3	&80.3/77.3	&84.1	&91.7	&68.3	&82.1 & 17.7x \\

    &large	        	&92.4/85.8	&83.9/80.8	&85.8	&92.2	&73.8	&85.1 & 3.8x\\
    &xlarge		&86.3/77.9	&73.8/70.5	&80.5	&87.8	&39.7	&76.7 & 1.0\\
\hline 
\multirow{4}{*}{ALBERT}  &base	        	    &89.3/82.1	&79.1/76.1	&81.9	&89.4	&63.5	&80.1 &21.1x \\
&large	        	    &90.9/84.1	&82.1/79.0	&83.8	&90.6	&68.4	&82.4 &6.5x\\
&xlarge		    &93.0/86.5	&85.9/83.1	&85.4	&91.9	&73.9	&85.5 & 2.4x\\
&xxlarge		& \textbf{94.1/88.3}	&\textbf{88.1/85.1}	&\textbf{88.0}	&\textbf{95.2}	&\textbf{82.3}	&\textbf{88.7} & 1.2x\\
\end{tabular}
\caption{Dev sets results for primary models pre-trained over BookCorpus and Wikipedia for 125k steps. The average score are averaged over the representative downstream tasks. For SQuAD datasets, the two metrics are first averaged.}
\label{table:overall_comparison}
\end{table}



We compare the performance of the primary models listed in Table \ref{table:overall_comparison} on the dev sets of several representative downstream tasks including SQuAD v1.1, SQuaAD v2.0, MNLI-m, SST-2 and RACE tasks. We first observe that BERT-xlarge gets significantly worse results than BERT-base on all metrics. This indicates that a model like BERT-xlarge is more difficult to train than those that have smaller parameter sizes. 
Another interesting observation we have is the speed of data throughput. The slowest one is the BERT-xlarge model, which we use as the baseline. ALBERT models have higher data throughput than their corresponding BERT models. As the model get larger, the differences between a BERT model and its light version become bigger. A ALBERT xlarge can be trained about 2.4x faster than BERT-x large. The 12-layer ALBERT-xxlarge model outperform all other models significantly. Compared to BERT-large, which has roughly 1.45x more parameters, ALBERT-xxlarge has an average of $3.6\%$ absolute improvements over 5 downstream tasks and $8.5\%$ absolute improvements over the RACE task. 
