%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Checklist}


\begin{enumerate}


\item For all authors...
\begin{enumerate}
  \item Do the main claims made in the abstract and introduction accurately reflect the paper's contributions and scope?
    \answerYes{}
  \item Did you describe the limitations of your work?
    \answerYes{We discussed the limitation at the end of Section 2 and in Section 3.2.}
  \item Did you discuss any potential negative societal impacts of your work?
    \answerYes{We discussed the reasoning paths we elicited from the models might not be faithful and future work should look at how to improve the factuality and how to better ground models' predictions.}
  \item Have you read the ethics review guidelines and ensured that your paper conforms to them?
    \answerYes{}
\end{enumerate}


\item If you are including theoretical results...
\begin{enumerate}
  \item Did you state the full set of assumptions of all theoretical results?
    \answerNA{This paper does not provide theoretical results.}
        \item Did you include complete proofs of all theoretical results?
   \answerNA{}
\end{enumerate}


\item If you ran experiments...
\begin{enumerate}
  \item Did you include the code, data, and instructions needed to reproduce the main experimental results (either in the supplemental material or as a URL)?
    \answerYes{We have provided the exact prompts used in the Appendix to reproduce the results. We also discussed the parameter settings in the experimental section.}
  \item Did you specify all the training details (e.g., data splits, hyperparameters, how they were chosen)?
    \answerYes{We included the exact dataset used (train, dev, or test) for evaluation. We also included all parameters used and how they are chosen for sampling the reasoning paths. We further provided an ablation study on results' sensitivity to parameter choices in Section 3.4.}
        \item Did you report error bars (e.g., with respect to the random seed after running experiments multiple times)?
    \answerYes{See Figure 3 in the main text and Figure 8 in Appendix.}
        \item Did you include the total amount of compute and the type of resources used (e.g., type of GPUs, internal cluster, or cloud provider)?
    \answerYes{See details in Appendix A.2.}
\end{enumerate}


\item If you are using existing assets (e.g., code, data, models) or curating/releasing new assets...
\begin{enumerate}
  \item If your work uses existing assets, did you cite the creators?
    \answerYes{We cited all the code/data/models used. There are two language models used that we anonymized based on the recommendation of the chairs. These models will be cited in the camera-ready version of the paper.}
  \item Did you mention the license of the assets?
    \answerYes{See Appendix A.4.}
  \item Did you include any new assets either in the supplemental material or as a URL?
    \answerNA{No new assets.}
  \item Did you discuss whether and how consent was obtained from people whose data you're using/curating?
    \answerNA{No human data is collected.}
  \item Did you discuss whether the data you are using/curating contains personally identifiable information or offensive content?
    \answerNA{All datasets used are public and no human data is collected.}
\end{enumerate}


\item If you used crowdsourcing or conducted research with human subjects...
\begin{enumerate}
  \item Did you include the full text of instructions given to participants and screenshots, if applicable?
    \answerNA{No crowdsourcing or research with human subjects.}
  \item Did you describe any potential participant risks, with links to Institutional Review Board (IRB) approvals, if applicable?
    \answerNA{}
  \item Did you include the estimated hourly wage paid to participants and the total amount spent on participant compensation?
    \answerNA{}
\end{enumerate}


\end{enumerate}